---
title: "Univariate analysis of heatwave time series biological data"
author: "Ariana S Huffmyer, E5 RoL Team"
edited by: "DM Becker-Polinski"
date: "20220830"
output: github_document
editor_options: 
  chunk_output_type: consoles
--- 

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")
if (!require("lsmeans")) install.packages("lsmeans")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(ggsignif)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(ggpubr)
library(broom)
library(multcompView)
library(cowplot)
library(datasets)
library(dplyr)
library(plyr)
library(grid)
library(gridExtra)
library(lsmeans)
```

# Load dataframe

Load in master dataframe generated from 1_assemble_data.Rmd.  
```{r}
master<-read.csv("gametogenesis_timeseries_analysis/Output/master_timeseries.csv")

# read in metadata to add patch to master timeseries
metadata <- read.csv("gametogenesis_timeseries_analysis/master_metadata.csv")

# add ACR- before each tag number
metadata <- metadata %>%
  mutate(colony_id = paste0("ACR-", colony_id))

#commbine master and metadata
master <- left_join(master, metadata)

#rename AUG september and october

master[master=="AUGUST"] <- "AUG"
master[master=="SEPTEMBER"] <- "SEP"
master[master=="OCTOBER"] <- "OCT"
master[master=="MARCH"] <- "MAR"
master[master=="APRIL"] <- "APR"
master[master=="MAY"] <- "MAY"
master[master=="JUNE"] <- "JUN"
master[master=="JULY"] <- "JUL"
master[master=="AUGUST"] <- "AUG"

```

# Univariate Responses  

Plot univariate response plots and then generate panel of all plots at the end of this section. Individual plots will not be displayed.  

### Plot: Symbiont Densities
 
View by treatment and timepoint.
```{r}

# Fit a mixed-effects model with 'patch' as a random effect
model <- lmer(cells.cm2 ~ timepoint + (1|patch), data = master)

# Perform ANOVA
anova_results <- Anova(model, type = "II")

# Post-hoc pairwise comparisons (optional)
pairwise_results <- emmeans(model, pairwise ~ timepoint, adjust = "tukey")


# Extract pairwise comparisons
pairwise_comparisons <- pairs(pairwise_results)

# Filter significant comparisons
significant_comparisons <- summary(pairwise_comparisons, test = "adjusted", by = NULL, infer = TRUE)

# Extract significant comparisons
significant_contrasts <- subset(significant_comparisons, p.value < 0.05)

# Create the plot
symbplot_cm2 <- master %>%
  filter(!is.na(cells.cm2)) %>%
  filter(!is.na(timepoint)) %>%
  select(colony_id, timepoint, cells.cm2) %>%
  
  ggplot(., aes(x = timepoint, y = cells.cm2, fill = timepoint)) +
  geom_boxplot(color = "black", fill = "white", alpha = 0.2) +
  geom_point(pch = 21, size = 2, position = position_jitterdodge(0.2)) +
  scale_x_discrete(limits = c("JAN", "FEB", "MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
  xlab("") + 
  ylab(expression(bold(paste("Symbiont Cells cm"^-2)))) +
  theme_classic() + 
  theme(legend.position = "none",
        axis.title = element_text(face = "bold", size = 20),
        axis.text = element_text(size = 12, color = "black"), 
        strip.text.x = element_text(face = "italic", size = 12))

# Add significance stars
symbplot_cm2 +
  geom_signif(data = significant_contrasts, 
              aes(x = timepoint, y = ymax + 5, annotations = as.character(p.value)), 
              inherit.aes = FALSE,
              manual = TRUE,
              step_increase = 0.1,
              vjust = -0.5)


```

### Plot: Host Protein   

View by treatment and timepoint.    
```{r}
host_prot_cm2<-master %>%
  filter(!is.na(host_prot_mg.cm2)) %>%
  filter(!is.na(timepoint))%>%
  select(colony_id, timepoint, host_prot_mg.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = host_prot_mg.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    ylim(0, 3.5)+
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") + 
    ylab(expression(bold(paste("Host Protein ( mg cm"^-2, ")"))))+
    theme_classic() + 
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40))


# Fit a linear model (ANOVA)
model <- aov(host_prot_mg.cm2 ~ timepoint, data = master)

# Perform ANOVA
anova_results <- Anova(model, type = "II")

# Display the ANOVA table
print(anova_results)

# Post-hoc pairwise comparisons (optional)
pairwise_results <- lsmeans(model, pairwise ~ timepoint, adjust = "tukey")

# Display the pairwise comparison results
print(pairwise_results)
```

### Plot: Holobiont protein  

View by treatment and timepoint.   
```{r}
holobiont_prot_cm2<-master %>%
  filter(!is.na(holobiont_prot_mg.cm2)) %>%
  filter(!is.na(timepoint))%>%
  select(colony_id, timepoint, holobiont_prot_mg.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = holobiont_prot_mg.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) +
    ylim(0, 3.5)+
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") +
    ylab(expression(bold(paste("Holobiont Protein (mg cm"^-2, ")"))))+
    theme_classic() + 
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40))


# Fit a linear model (ANOVA)
model <- aov(holobiont_prot_mg.cm2 ~ timepoint, data = master)

# Perform ANOVA
anova_results <- Anova(model, type = "II")

# Display the ANOVA table
print(anova_results)

# Post-hoc pairwise comparisons (optional)
pairwise_results <- lsmeans(model, pairwise ~ timepoint, adjust = "tukey")

# Display the pairwise comparison results
print(pairwise_results)


```


### Plot: CHL-a

View by treatment and timepoint.   
```{r}
chla_cm2<-master %>%
  filter(!is.na(chla.ug.cm2)) %>%
  filter(!is.na(timepoint))%>%
  select(colony_id, timepoint, chla.ug.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = chla.ug.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) +
    ylim(0, 12)+
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") + 
    ylab(expression(bold(paste("Chlorophyll a (", mu, "g cm"^-2, ")"))))+
    theme_classic() + 
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40))


# Fit a linear model (ANOVA)
model <- aov(chla.ug.cm2 ~ timepoint, data = master)

# Perform ANOVA
anova_results <- Anova(model, type = "II")

# Display the ANOVA table
print(anova_results)

# Post-hoc pairwise comparisons (optional)
pairwise_results <- lsmeans(model, pairwise ~ timepoint, adjust = "tukey")

# Display the pairwise comparison results
print(pairwise_results)

```

### Plot: CHL-c

View by treatment and timepoint.   
```{r}
chlc_prot_cm2<-master %>%
  filter(!is.na(chlc2.ug.cm2)) %>%
  filter(!is.na(timepoint))%>%
  select(colony_id, timepoint, chlc2.ug.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = chlc2.ug.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) +
    ylim(0, 12)+
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") + 
    ylab(expression(bold(paste("Chlorophyll c2 (", mu, "g cm"^-2, ")"))))+
    theme_classic() + 
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40))

# Fit a linear model (ANOVA)
model <- aov(chlc2.ug.cm2 ~ timepoint, data = master)

# Perform ANOVA
anova_results <- Anova(model, type = "II")

# Display the ANOVA table
print(anova_results)

# Post-hoc pairwise comparisons (optional)
pairwise_results <- lsmeans(model, pairwise ~ timepoint, adjust = "tukey")

# Display the pairwise comparison results
print(pairwise_results)
```


Join all plots for each normalization together.  

```{r}
phys_sym_figure<- plot_grid(symbplot_cm2, chla_cm2, chlc_prot_cm2, ncol=3, nrow=1, labels = c('A', 'B', 'C'),label_size = 80,rel_heights= 1, rel_widths = 1, label_y=1, align="h") 

#add x label to plot_GRID
x.grob <- textGrob("TIMEPOINT", 
                   gp=gpar(fontface="bold", col="black", fontsize=50))

#add to plot
phys_sym_figure <- grid.arrange(arrangeGrob(phys_sym_figure, bottom = x.grob))

ggsave(filename="gametogenesis_timeseries_analysis/Figures/Phys_Sym_Figure.pdf", plot=phys_sym_figure, dpi=300, width=49, height=20, units="in")
ggsave(filename="gametogenesis_timeseries_analysis/Figures/Phys_Sym_Figure.png", plot=phys_sym_figure, dpi=300, width=49, height=20, units="in")

phys_host_figure<-plot_grid( host_prot_cm2, holobiont_prot_cm2, ncol=2, nrow=1, labels = c('A', 'B'),label_size = 80,rel_heights= 1, rel_widths = 1, label_y=1, align="h")

#add x label to plot_GRID
x.grob <- textGrob("TIMEPOINT", 
                   gp=gpar(fontface="bold", col="black", fontsize=50))

#add to plot
phys_host_figure <- grid.arrange(arrangeGrob(phys_host_figure, bottom = x.grob))

ggsave(filename="gametogenesis_timeseries_analysis/Figures/Phys_Host_Figure.pdf", plot=phys_host_figure, dpi=300, width=35, height=20, units="in")
ggsave(filename="gametogenesis_timeseries_analysis/Figures/Phys_Host_Figure.png", plot=phys_host_figure, dpi=300, width=35, height=20, units="in")
```

