---
title: "Univariate analysis of heatwave time series biological data"
author: "Ariana S Huffmyer, E5 RoL Team"
edited by: "DM Becker-Polinski"
date: "20220830"
output: github_document
editor_options: 
  chunk_output_type: consoles
--- 

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")
if (!require("lsmeans")) install.packages("lsmeans")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(ggsignif)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(ggpubr)
library(broom)
library(multcomp)
library(multcompView)
library(cowplot)
library(datasets)
library(dplyr)
library(plyr)
library(grid)
library(gridExtra)
library(lsmeans)
```

# Load dataframe

Load in master dataframe generated from 1_assemble_data.Rmd.  
```{r}
master<-read.csv("gametogenesis_timeseries/gametogenesis_timeseries_analysis/Output/master_timeseries.csv")

# read in metadata to add patch to master timeseries
metadata <- read.csv("gametogenesis_timeseries/gametogenesis_timeseries_analysis/master_metadata.csv")

# add ACR- before each tag number
metadata <- metadata %>%
  mutate(colony_id = paste0("ACR-", colony_id))

#commbine master and metadata
master <- left_join(master, metadata)

#rename AUG september and october

master[master=="AUGUST"] <- "AUG"
master[master=="SEPTEMBER"] <- "SEP"
master[master=="OCTOBER"] <- "OCT"
master[master=="MARCH"] <- "MAR"
master[master=="APRIL"] <- "APR"
master[master=="MAY"] <- "MAY"
master[master=="JUNE"] <- "JUN"
master[master=="JULY"] <- "JUL"
master[master=="AUGUST"] <- "AUG"
master[master=="JANUARY"] <- "JAN"
master[master=="FEBRUARY"] <- "FEB"


```

# Univariate Responses  

Plot univariate response plots and then generate panel of all plots at the end of this section. Individual plots will not be displayed.  

### Plot: Symbiont Densities
 
View by treatment and timepoint.
```{r}
# Ensure timepoint is a factor
master$timepoint <- factor(master$timepoint, levels = c("JAN", "FEB", "MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT", "NOV", "DEC"))

# Fit a mixed-effects model with 'patch' as a random effect
model.sym <- lmer(cells.cm2 ~ timepoint + (1|patch), data = master)
qqPlot(residuals(model.sym))

# Residuals are not normally distributed. Attempt with log transformation.   
sym_model<-lmer(log(cells.cm2)~timepoint+(1|patch), data=master)
qqPlot(residuals(sym_model))

# Identify the outliers
outliers <- c(355,94,194,364, 214, 48)

# remove outliers
data_clean_sym <- master[-outliers, ]

# Residuals are not normally distributed. Attempt with log transformation.  
# Fit the model with January as the intercept
sym_model <- lmer(log(cells.cm2) ~ timepoint + (1 | patch), data = data_clean_sym)

# Check residuals for normality
qqPlot(residuals(sym_model))

# Get the summary of the model
summary(sym_model)

# Post-hoc pairwise comparisons
pairwise_results.sym <- emmeans(sym_model, pairwise ~ timepoint, adjust = "tukey")

# Generate letters for significant differences
timepoint_cld_sym <- cld(pairwise_results.sym, Letters = letters)

# Remove outlier data
outlier_value <- 2430445.8
filtered_sym <- master %>%
  filter(round(cells.cm2, digits = 1) != round(outlier_value, digits = 1)) %>%
  filter(!is.na(cells.cm2)) %>%
  filter(!is.na(timepoint))

# Get the maximum value of cells.cm2
max_cells_cm2 <- max(filtered_sym$cells.cm2, na.rm = TRUE)

# Create the plot
symbplot_cm2 <- filtered_sym %>%
  ggplot(aes(x = timepoint, y = cells.cm2, fill = timepoint)) +
  geom_boxplot(color = "black", fill = "white", alpha = 0.2) +
  geom_point(pch = 21, size = 2, position = position_jitterdodge(0.2)) +
  geom_text(data = timepoint_cld_sym, aes(label = .group, x = timepoint, y = 0.96 * max_cells_cm2), vjust = -0.5, size = 14, color = "black") +
 # scale_x_discrete(limits = c("JAN", "FEB", "MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
  xlab("") + 
  ylab(expression(bold(paste("Symbiont Cells cm"^-2)))) +
  theme_classic() + 
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold", size = 60),
    axis.text = element_text(size = 40, color = "black"),
    strip.text.x = element_text(face = "italic", size = 40));symbplot_cm2

```


### Plot: CHL-a

View by timepoint  
```{r}

# Fit a mixed-effects model with 'patch' as a random effect, check normality
model.chla <- lmer(chla.ug.cm2 ~ timepoint + (1|patch), data = master)
qqPlot(residuals(model.chla))

# Residuals are not normally distributed. Attempt with log transformation.  
chla_model<-lmer(log(chla.ug.cm2)~timepoint+(1|patch), data=master)
qqPlot(residuals(chla_model))

# Identify the outliers
outliers <- c(94,308, 45, 346)

# remove outliers
data_clean_chla <- master[-outliers, ]

# Residuals are not normally distributed. Attempt with log transformation.  
chla_model<-lmer(log(chla.ug.cm2)~timepoint+(1|patch), data=data_clean_chla)
qqPlot(residuals(chla_model))

# Get the summary of the model
summary(chla_model)

# Perform Type III ANOVA
anova_result <- Anova(chla_model, type = 3)

# Post-hoc pairwise comparisons
pairwise_results.chla <- emmeans(chla_model, pairwise ~ timepoint, adjust = "tukey")

# Generate letters for significant differences
timepoint_cld_chla <- cld(pairwise_results.chla, Letters = letters)

# Define a threshold for identifying outliers for the 'Value' column
threshold <- 10  # For example, we consider values greater than 50 as outliers

# Filter out values beyond the threshold for the 'Value' column
filtered_chla <- master[master$chla.ug.cm2 <= threshold, ]

# Get the maximum value of cells.cm2
max_chla <- max(master$chla.ug.cm2, na.rm = TRUE)  # Make sure to handle any missing values

# make chla plot
chla_cm2 <- filtered_chla %>%
  filter(!is.na(chla.ug.cm2)) %>%
  filter(!is.na(timepoint))%>%
  dplyr::select(colony_id, timepoint, chla.ug.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = chla.ug.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) +
    ylim(0, 12)+
  geom_text(data = timepoint_cld_chla, aes(label = .group, x = timepoint, y = 1.0*max_chla), vjust = -0.5, size = 14, color = "black") +
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") + 
    ylab(expression(bold(paste("Chlorophyll a (", mu, "g cm"^-2, ")"))))+
    theme_classic() + 
    scale_y_continuous(limits = c(0, 8))  + # Adjust limits as needed
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40)); chla_cm2

```

### Plot: CHL-c

View by timepoint  
```{r}
# Fit a mixed-effects model with 'patch' as a random effect
model.chlc <- lmer(chlc2.ug.cm2 ~ timepoint + (1|patch), data = master)
qqPlot(residuals(chla_model))

# Get the summary of the model
summary(chlc_model)

# Perform Type III ANOVA
anova_result <- Anova(chlc_model, type = 3)

# Post-hoc pairwise comparisons
pairwise_results.chlc <- emmeans(chlc_model, pairwise ~ timepoint, adjust = "tukey")

# Generate letters for significant differences
timepoint_cld_chlc <- cld(pairwise_results.chlc, Letters = letters)

# Get the maximum value of cells.cm2
max_chlc <- max(master$chlc2.ug.cm2, na.rm = TRUE)  # Make sure to handle any missing values

# Make chlc plot
chlc_prot_cm2 <- master %>%
  filter(!is.na(chlc2.ug.cm2)) %>%
  filter(!is.na(timepoint))%>%
  dplyr::select(colony_id, timepoint, chlc2.ug.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = chlc2.ug.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) +
    ylim(0, 12)+
    geom_text(data = timepoint_cld_chlc, aes(label = .group, x = timepoint, y = 1.0*max_chlc), vjust = -0.5, size = 14, color = "black") +
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") + 
    ylab(expression(bold(paste("Chlorophyll c2 (", mu, "g cm"^-2, ")"))))+
    theme_classic() + 
    scale_y_continuous(limits = c(0, 4.5))  + # Adjust limits as needed
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40)) ; chlc_prot_cm2


```



### Plot: total-chl

View by timepoint  
```{r}
# Fit a mixed-effects model with 'patch' as a random effect
model.totchl <- lmer(Total_Chl_cm2 ~ timepoint + (1|patch), data = master)
qqPlot(residuals(model.totchl))

# Residuals are not normally distributed. Attempt with log transformation.  
totchl_model<-lmer(log(Total_Chl_cm2)~timepoint+(1|patch), data=master)
qqPlot(residuals(totchl_model))

# Identify the outliers
outliers <- c(94,308)

# remove outliers
data_clean_tot <- master[-outliers, ]

# Residuals are not normally distributed. Attempt with log transformation.  
totchl_model<-lmer(log(Total_Chl_cm2)~timepoint+(1|patch), data=data_clean_tot)
qqPlot(residuals(totchl_model))

# Get the summary of the model
summary(totchl_model)

# Perform Type III ANOVA
anova_result <- Anova(totchl_model, type = 3)

# Post-hoc pairwise comparisons
pairwise_results.tot <- emmeans(totchl_model, pairwise ~ timepoint, adjust = "tukey")

# Generate letters for significant differences
timepoint_cld_tot <- cld(pairwise_results.tot, Letters = letters)

# Get the maximum value of cells.cm2
max_tot <- max(master$Total_Chl_cm2, na.rm = TRUE)  # Make sure to handle any missing values

# Make chlc plot
tot.chl <- master %>%
  filter(!is.na(Total_Chl_cm2)) %>%
  filter(!is.na(timepoint))%>%
  dplyr::select(colony_id, timepoint, Total_Chl_cm2)%>%
  
  ggplot(., aes(x = timepoint, y = Total_Chl_cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) +
    ylim(0, 12)+
      geom_text(data = timepoint_cld_tot, aes(label = .group, x = timepoint, y = 1.0*max_tot), vjust = -0.5, size = 14, color = "black") +
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") + 
    ylab(expression(bold(paste("Total Chlorophyll (", mu, "g cm"^-2, ")"))))+
    theme_classic() + 
    scale_y_continuous(limits = c(0, 10))  + # Adjust limits as needed
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40)) ; tot.chl


```

### Plot: Sym biomass 

View by timepoint  
```{r}
# Fit a mixed-effects model with 'patch' as a random effect
model.sym.bio <- lmer(Sym_AFDW.mg.cm2 ~ timepoint + (1|patch), data = master)
qqPlot(residuals(model.sym.bio))

# Residuals are not normally distributed. Attempt with log transformation.  
sym.bio.model<-lmer(log(Sym_AFDW.mg.cm2)~timepoint+(1|patch), data=master)
qqPlot(residuals(sym.bio.model))

# Identify the outliers
outliers <- c(282,281,259,23)

# remove outliers
data_clean_sym.bio <- master[-outliers, ]

# Residuals are not normally distributed. Attempt with log transformation.  
sym.bio.model<-lmer(log(Sym_AFDW.mg.cm2)~timepoint+(1|patch), data=data_clean_sym.bio)
qqPlot(residuals(sym.bio.model))

# Get the summary of the model
summary(sym.bio.model)

# Perform Type III ANOVA
anova_result <- Anova(sym.bio.model, type = 3)

# Post-hoc pairwise comparisons
pairwise_results.sym.bio <- emmeans(sym.bio.model, pairwise ~ timepoint, adjust = "tukey")

# Generate letters for significant differences
timepoint_sym_bio <- cld(pairwise_results.sym.bio, Letters = letters)

# Define a threshold for identifying outliers for the 'Value' column
threshold <- 0  # For example, we consider values greater than 50 as outliers

# Filter out values beyond the threshold for the 'Value' column
filtered_sym_bio <- master[master$Sym_AFDW.mg.cm2 >= threshold, ]

# Get the maximum value of cells.cm2
max_sym_bio <- max(filtered_sym_bio$Sym_AFDW.mg.cm2, na.rm = TRUE)  # Make sure to handle any missing values

# Make holobiont protein plot
sym_bio_cm2 <- filtered_sym_bio %>%
  filter(!is.na(Sym_AFDW.mg.cm2)) %>%
  filter(!is.na(timepoint))%>%
  dplyr::select(colony_id, timepoint, Sym_AFDW.mg.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = Sym_AFDW.mg.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) +
    ylim(0, 3)+
    geom_text(data = timepoint_host_bio, aes(label = .group, x = timepoint, y = 1.2*max_sym_bio), vjust = -0.5, size = 14, color = "black") +
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") +
    ylab(expression(bold(paste("Symbiont Biomass (mg cm"^-2, ")"))))+
    theme_classic() + 
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40)); sym_bio_cm2


```


### Plot: Host Protein   

View by timepoint   
```{r}
# Fit a mixed-effects model with 'patch' as a random effect
model.host.prot <- lmer(host_prot_mg.cm2 ~ timepoint + (1|patch), data = master)
qqPlot(residuals(model.host.prot))

# Residuals are not normally distributed. Attempt with log transformation.  
host.prot.model<-lmer(log(host_prot_mg.cm2)~timepoint+(1|patch), data=master)
qqPlot(residuals(host.prot.model))

# Identify the outliers
outliers <- c(259,51, 45, 143)

# remove outliers
data_clean_host_prot <- master[-outliers, ]

# Residuals are not normally distributed. Attempt with log transformation.  
host.prot.model<-lmer(log(host_prot_mg.cm2)~timepoint+(1|patch), data=data_clean_host_prot)
qqPlot(residuals(host.prot.model))

# Get the summary of the model
summary(host.prot.model)

# Perform Type III ANOVA
anova_result <- Anova(host.prot.model, type = 3)

# Post-hoc pairwise comparisons
pairwise_results.host.prot <- emmeans(host.prot.model, pairwise ~ timepoint, adjust = "tukey")

# Generate letters for significant differences
timepoint_host_prot <- cld(pairwise_results.host.prot, Letters = letters)

# Get the maximum value of cells.cm2
max_host_prot <- max(master$host_prot_mg.cm2, na.rm = TRUE)  # Make sure to handle any missing values

# Make plot for host protein
host_prot_cm2 <- master %>%
  filter(!is.na(host_prot_mg.cm2)) %>%
  filter(!is.na(timepoint))%>%
  dplyr::select(colony_id, timepoint, host_prot_mg.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = host_prot_mg.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    ylim(0, 3.5)+
        geom_text(data = timepoint_host_prot, aes(label = .group, x = timepoint, y = 1.0*max_host_prot), vjust = -0.5, size = 14, color = "black") +
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") + 
    ylab(expression(bold(paste("Host Protein ( mg cm"^-2, ")"))))+
    theme_classic() + 
    scale_y_continuous(limits = c(0, 2.5))  + # Adjust limits as needed
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40)); host_prot_cm2


```


### Plot: Holobiont protein  

View by timepoint  
```{r}
# Fit a mixed-effects model with 'patch' as a random effect
model.holo.prot <- lmer(holobiont_prot_mg.cm2 ~ timepoint + (1|patch), data = master)
qqPlot(residuals(model.holo.prot))

# Residuals are not normally distributed. Attempt with log transformation.  
holo.prot.model<-lmer(log(holobiont_prot_mg.cm2)~timepoint+(1|patch), data=master)
qqPlot(residuals(holo.prot.model))

# Identify the outliers
outliers <- c(18,1, 259,7,6,201)

# remove outliers
data_clean_holo_prot <- master[-outliers, ]

# Residuals are not normally distributed. Attempt with log transformation.  
holo.prot.model<-lmer(log(holobiont_prot_mg.cm2)~timepoint+(1|patch), data=data_clean_holo_prot)
qqPlot(residuals(holo.prot.model))

# Get the summary of the model
summary(holo.prot.model)

# Perform Type III ANOVA
anova_result <- Anova(holo.prot.model, type = 3)

# Post-hoc pairwise comparisons
pairwise_results.holo.prot <- emmeans(holo.prot.model, pairwise ~ timepoint, adjust = "tukey")

# Generate letters for significant differences
timepoint_holo_prot <- cld(pairwise_results.holo.prot, Letters = letters)

# Get the maximum value of cells.cm2
max_holo_prot <- max(master$holobiont_prot_mg.cm2, na.rm = TRUE)  # Make sure to handle any missing values

# Make holobiont protein plot
holobiont_prot_cm2 <- master %>%
  filter(!is.na(holobiont_prot_mg.cm2)) %>%
  filter(!is.na(timepoint))%>%
  dplyr::select(colony_id, timepoint, holobiont_prot_mg.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = holobiont_prot_mg.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) +
    ylim(0, 3.5)+
          geom_text(data = timepoint_holo_prot, aes(label = .group, x = timepoint, y = 1.0*max_holo_prot), vjust = -0.5, size = 14, color = "black") +
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") +
    ylab(expression(bold(paste("Holobiont Protein (mg cm"^-2, ")"))))+
    theme_classic() + 
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40)); holobiont_prot_cm2



```


### Plot: Host biomass 

View by timepoint  
```{r}
# Fit a mixed-effects model with 'patch' as a random effect
model.host.bio <- lmer(Host_AFDW.mg.cm2 ~ timepoint + (1|patch), data = master)
qqPlot(residuals(model.host.bio))

# Residuals are not normally distributed. Attempt with log transformation.  
host.bio.model<-lmer(log(Host_AFDW.mg.cm2)~timepoint+(1|patch), data=master)
qqPlot(residuals(host.bio.model))

# Identify the outliers
outliers <- c(295,293,6, 271,9)

# remove outliers
data_clean_host.bio <- master[-outliers, ]

# Residuals are not normally distributed. Attempt with log transformation.  
host.bio.model<-lmer(log(Host_AFDW.mg.cm2)~timepoint+(1|patch), data=data_clean_host.bio)
qqPlot(residuals(host.bio.model))

# Get the summary of the model
summary(host.bio.model)

# Perform Type III ANOVA
anova_result <- Anova(host.bio.model, type = 3)

# Post-hoc pairwise comparisons
pairwise_results.host.bio <- emmeans(host.bio.model, pairwise ~ timepoint, adjust = "tukey")

# Generate letters for significant differences
timepoint_host_bio <- cld(pairwise_results.host.bio, Letters = letters)

# Define a threshold for identifying outliers for the 'Value' column
threshold <- 9  # For example, we consider values greater than 50 as outliers

# Filter out values beyond the threshold for the 'Value' column
filtered_host_bio <- master[master$Host_AFDW.mg.cm2 <= threshold, ]

# Get the maximum value of cells.cm2
max_host_bio <- max(filtered_host_bio$Host_AFDW.mg.cm2, na.rm = TRUE)  # Make sure to handle any missing values

# Make holobiont protein plot
host_bio_cm2 <- filtered_host_bio %>%
  filter(!is.na(Host_AFDW.mg.cm2)) %>%
  filter(!is.na(timepoint))%>%
  dplyr::select(colony_id, timepoint, Host_AFDW.mg.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = Host_AFDW.mg.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) +
    ylim(0, 6)+
    geom_text(data = timepoint_host_bio, aes(label = .group, x = timepoint, y = 1.0*max_host_bio), vjust = -0.5, size = 14, color = "black") +
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") +
    ylab(expression(bold(paste("Host Biomass (mg cm"^-2, ")"))))+
    theme_classic() + 
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40)); host_bio_cm2


```


Join all plots for each normalization together.  

```{r}
phys_sym_figure<- plot_grid(symbplot_cm2, sym_bio_cm2, tot.chl, ncol=3, nrow=1, labels = c('A', 'B', 'C'),label_size = 80,rel_heights= 1, rel_widths = 1, label_y=1, align="h") 

#add x label to plot_GRID
x.grob <- textGrob("TIMEPOINT", 
                   gp=gpar(fontface="bold", col="black", fontsize=50))

#add to plot
phys_sym_figure <- grid.arrange(arrangeGrob(phys_sym_figure, bottom = x.grob))

ggsave(filename="gametogenesis_timeseries/gametogenesis_timeseries_analysis/l;'
       Figures/Phys_Sym_Figure.png", plot=phys_sym_figure, dpi=300, width=49, height=20, units="in")

phys_host_figure<-plot_grid(host_prot_cm2, holobiont_prot_cm2, host_bio_cm2, ncol=3, nrow=1, labels = c('A', 'B', 'C'),label_size = 80,rel_heights= 1, rel_widths = 1, label_y=1, align="h")

#add x label to plot_GRID
x.grob <- textGrob("TIMEPOINT", 
                   gp=gpar(fontface="bold", col="black", fontsize=50))

#add to plot
phys_host_figure <- grid.arrange(arrangeGrob(phys_host_figure, bottom = x.grob))

ggsave(filename="gametogenesis_timeseries/gametogenesis_timeseries_analysis/Figures/Phys_Host_Figure.png", plot=phys_host_figure, dpi=300, width=45, height=20, units="in")
```

