---
title: "Univariate analysis of heatwave time series biological data"
author: "Ariana S Huffmyer, E5 RoL Team"
edited by: "DM Becker-Polinski"
date: "20220830"
output: github_document
editor_options: 
  chunk_output_type: consoles
--- 

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")
if (!require("lsmeans")) install.packages("lsmeans")

# load packages
library(tidyverse)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(ggsignif)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(corrr)
library(ggcorrplot)
library(GGally)
library(ggpubr)
library(broom)
library(multcomp)
library(multcompView)
library(cowplot)
library(datasets)
library(dplyr)
library(plyr)
library(grid)
library(gridExtra)
library(lsmeans)
```

# Load dataframe

Load in master dataframe generated from 1_assemble_data.Rmd.  
```{r}
master<-read.csv("gametogenesis_timeseries_analysis/Output/master_timeseries.csv")

# read in metadata to add patch to master timeseries
metadata <- read.csv("gametogenesis_timeseries_analysis/master_metadata.csv")

# add ACR- before each tag number
metadata <- metadata %>%
  mutate(colony_id = paste0("ACR-", colony_id))

#commbine master and metadata
master <- left_join(master, metadata)

#rename AUG september and october

master[master=="AUGUST"] <- "AUG"
master[master=="SEPTEMBER"] <- "SEP"
master[master=="OCTOBER"] <- "OCT"
master[master=="MARCH"] <- "MAR"
master[master=="APRIL"] <- "APR"
master[master=="MAY"] <- "MAY"
master[master=="JUNE"] <- "JUN"
master[master=="JULY"] <- "JUL"
master[master=="AUGUST"] <- "AUG"
master[master=="JANUARY"] <- "JAN"
master[master=="FEBRUARY"] <- "FEB"


```

# Univariate Responses  

Plot univariate response plots and then generate panel of all plots at the end of this section. Individual plots will not be displayed.  

### Plot: Symbiont Densities
 
View by treatment and timepoint.
```{r}
# Ensure timepoint is a factor
master$timepoint <- factor(master$timepoint, levels = c("JAN", "FEB", "MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT", "NOV", "DEC"))

# Fit a mixed-effects model with 'patch' as a random effect
model.sym <- lmer(cells.cm2 ~ timepoint + (1|patch), data = master)
qqPlot(residuals(model.sym))

# Residuals are not normally distributed. Attempt with log transformation.   
sym_model<-lmer(log(cells.cm2)~timepoint+(1|patch), data=master)
qqPlot(residuals(sym_model))

# Identify the outliers
outliers <- c(355,94,194,364, 214, 48)

# remove outliers
data_clean_sym <- master[-outliers, ]

# Residuals are not normally distributed. Attempt with log transformation.  
# Fit the model with January as the intercept
sym_model <- lmer(log(cells.cm2) ~ timepoint + (1 | patch), data = data_clean_sym)

# Check residuals for normality
qqPlot(residuals(sym_model))

# Get the summary of the model
summary(sym_model)

# Post-hoc pairwise comparisons
pairwise_results.sym <- emmeans(sym_model, pairwise ~ timepoint, adjust = "tukey")

# Generate letters for significant differences
timepoint_cld <- cld(pairwise_results.sym, Letters = letters)

# Remove outlier data
outlier_value <- 2430445.8
filtered_sym <- master %>%
  filter(round(cells.cm2, digits = 1) != round(outlier_value, digits = 1)) %>%
  filter(!is.na(cells.cm2)) %>%
  filter(!is.na(timepoint))

# Get the maximum value of cells.cm2
max_cells_cm2 <- max(filtered_sym$cells.cm2, na.rm = TRUE)

# Create the plot
symbplot_cm2 <- filtered_sym %>%
  ggplot(aes(x = timepoint, y = cells.cm2, fill = timepoint)) +
  geom_boxplot(color = "black", fill = "white", alpha = 0.2) +
  geom_point(pch = 21, size = 2, position = position_jitterdodge(0.2)) +
  geom_text(data = timepoint_cld, aes(label = .group, x = timepoint, y = 0.96 * max_cells_cm2), vjust = -0.5, size = 6, color = "black") +
 # scale_x_discrete(limits = c("JAN", "FEB", "MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
  xlab("") + 
  ylab(expression(bold(paste("Symbiont Cells cm"^-2)))) +
  theme_classic() + 
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold", size = 60),
    axis.text = element_text(size = 40, color = "black"),
    strip.text.x = element_text(face = "italic", size = 40));symbplot_cm2

```


### Plot: CHL-a

View by treatment and timepoint.   
```{r}

# Fit a mixed-effects model with 'patch' as a random effect, check normality
model.chla <- lmer(chla.ug.cm2 ~ timepoint + (1|patch), data = master)
qqPlot(residuals(model.chla))

# Residuals are not normally distributed. Attempt with log transformation.  
chla_model<-lmer(log(chla.ug.cm2)~timepoint+(1|patch), data=master)
qqPlot(residuals(chla_model))

# Identify the outliers
outliers <- c(94,308)

# remove outliers
data_clean_chla <- master[-outliers, ]

# Residuals are not normally distributed. Attempt with log transformation.  
chla_model<-lmer(log(chla.ug.cm2)~timepoint+(1|patch), data=data_clean_chla)
qqPlot(residuals(chla_model))

# Get the summary of the model
summary(chla_model)

# Perform Type III ANOVA
anova_result <- Anova(chla_model, type = 3)

# Post-hoc pairwise comparisons (optional)
pairwise_results.chla <- emmeans(chla_model, pairwise ~ timepoint, adjust = "tukey")

# Extract pairwise comparisons
pairwise_comparisons.chla <- pairs(pairwise_results.chla)

# Filter significant comparisons
significant_comparisons.chla <- summary(pairwise_comparisons.chla, test = "adjusted", by = NULL, infer = TRUE)

# Extract significant comparisons
significant_contrasts.chla <- subset(significant_comparisons.chla, p.value < 0.05)

# Define a threshold for identifying outliers for the 'Value' column
threshold <- 10  # For example, we consider values greater than 50 as outliers

# Filter out values beyond the threshold for the 'Value' column
filtered_master <- master[master$chla.ug.cm2 <= threshold, ]

# Create a data frame with significant timepoints
significant_timepoints_chla <- c("JAN", "FEB", "JUL", "JUN", "APR", "OCT", "SEPT")

# Get the maximum value of cells.cm2
max_chla <- max(master$chla.ug.cm2, na.rm = TRUE)  # Make sure to handle any missing values

# make chla plot
chla_cm2 <- filtered_master %>%
  filter(!is.na(chla.ug.cm2)) %>%
  filter(!is.na(timepoint))%>%
  select(colony_id, timepoint, chla.ug.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = chla.ug.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) +
    ylim(0, 12)+
    geom_text(data = data.frame(timepoint = significant_timepoints_chla), aes(label = ifelse(timepoint == "JUL", "**", "***")), x = significant_timepoints_chla, y = 0.1 + max_chla, vjust = -0.5, size = ifelse(significant_timepoints_chla == "JUL", 14, 14), color = "black") +  # Add significance indicators
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") + 
    ylab(expression(bold(paste("Chlorophyll a (", mu, "g cm"^-2, ")"))))+
    theme_classic() + 
    scale_y_continuous(limits = c(0, 8))  + # Adjust limits as needed
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40)); chla_cm2

```

### Plot: CHL-c

View by treatment and timepoint.   
```{r}

# Check for normality
qqnorm(master$chlc2.ug.cm2)

# Fit a mixed-effects model with 'patch' as a random effect
model.chlc <- lmer(chlc2.ug.cm2 ~ timepoint + (1|patch), data = master)

# Get the summary of the model
summary(model.chlc)

# Perform Type III ANOVA
anova_result <- Anova(model.chlc, type = 3)

# Post-hoc pairwise comparisons (optional)
pairwise_results.chlc <- emmeans(model.chlc, pairwise ~ timepoint, adjust = "tukey")

# Extract pairwise comparisons
pairwise_comparisons.chlc <- pairs(pairwise_results.chlc)

# Filter significant comparisons
significant_comparisons.chlc <- summary(pairwise_comparisons.chlc, test = "adjusted", by = NULL, infer = TRUE)

# Extract significant comparisons
significant_contrasts.chlc <- subset(significant_comparisons.chlc, p.value < 0.05)

# Make chlc plot
chlc_prot_cm2 <- master %>%
  filter(!is.na(chlc2.ug.cm2)) %>%
  filter(!is.na(timepoint))%>%
  select(colony_id, timepoint, chlc2.ug.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = chlc2.ug.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) +
    ylim(0, 12)+
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") + 
    ylab(expression(bold(paste("Chlorophyll c2 (", mu, "g cm"^-2, ")"))))+
    theme_classic() + 
    scale_y_continuous(limits = c(0, 4.5))  + # Adjust limits as needed
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40)) ; chlc_prot_cm2


```

### Plot: Host Protein   

View by treatment and timepoint.    
```{r}

# check for normality
qqnorm(master$host_prot_mg.cm2)

# Fit a mixed-effects model with 'patch' as a random effect
model.host.prot <- lmer(host_prot_mg.cm2 ~ timepoint + (1|patch), data = master)

# Get the summary of the model
summary(model.host.prot)

# Perform Type III ANOVA
anova_result <- Anova(model.host.prot, type = 3)

# Post-hoc pairwise comparisons (optional)
pairwise_results.host.prot <- emmeans(model.host.prot, pairwise ~ timepoint, adjust = "tukey")

# Extract pairwise comparisons
pairwise_comparisons.host.prot <- pairs(pairwise_results.host.prot)

# Filter significant comparisons
significant_comparisons.host.prot <- summary(pairwise_comparisons.host.prot, test = "adjusted", by = NULL, infer = TRUE)

# Extract significant comparisons
significant_contrasts.host.prot <- subset(significant_comparisons.host.prot, p.value < 0.05)

# Make plot for host protein
host_prot_cm2 <- master %>%
  filter(!is.na(host_prot_mg.cm2)) %>%
  filter(!is.na(timepoint))%>%
  select(colony_id, timepoint, host_prot_mg.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = host_prot_mg.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) + 
    ylim(0, 3.5)+
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") + 
    ylab(expression(bold(paste("Host Protein ( mg cm"^-2, ")"))))+
    theme_classic() + 
    scale_y_continuous(limits = c(0, 2.5))  + # Adjust limits as needed
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40)); host_prot_cm2


```


### Plot: Holobiont protein  

View by treatment and timepoint.   
```{r}

# check for normality
qqnorm(master$holobiont_prot_mg.cm2)

# Fit a mixed-effects model with 'patch' as a random effect
model.holo.prot <- lmer(holobiont_prot_mg.cm2 ~ timepoint + (1|patch), data = master)

# Get the summary of the model
summary(model.holo.prot)

# Perform Type III ANOVA
anova_result <- Anova(model.holo.prot, type = 3)

# Post-hoc pairwise comparisons (optional)
pairwise_results.holo.prot <- emmeans(model.holo.prot, pairwise ~ timepoint, adjust = "tukey")

# Extract pairwise comparisons
pairwise_comparisons.holo.prot <- pairs(pairwise_results.holo.prot)

# Filter significant comparisons
significant_comparisons.holo.prot <- summary(pairwise_comparisons.holo.prot, test = "adjusted", by = NULL, infer = TRUE)

# Extract significant comparisons
significant_contrasts.holo.prot <- subset(significant_comparisons.holo.prot, p.value < 0.05)

# Make holobiont protein plot
holobiont_prot_cm2 <- master %>%
  filter(!is.na(holobiont_prot_mg.cm2)) %>%
  filter(!is.na(timepoint))%>%
  select(colony_id, timepoint, holobiont_prot_mg.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = holobiont_prot_mg.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) +
    ylim(0, 3.5)+
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") +
    ylab(expression(bold(paste("Holobiont Protein (mg cm"^-2, ")"))))+
    theme_classic() + 
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40)); holobiont_prot_cm2



```


### Plot: Host biomass 

View by treatment and timepoint.   
```{r}

# check for normality
qqnorm(master$Host_AFDW.mg.cm2)

# Fit a mixed-effects model with 'patch' as a random effect
model.host.bio <- lmer(Host_AFDW.mg.cm2 ~ timepoint + (1|patch), data = master)

# Get the summary of the model
summary(model.host.bio)

# Perform Type III ANOVA
anova_result <- Anova(model.host.bio, type = 3)

# Post-hoc pairwise comparisons (optional)
pairwise_results.host.bio <- emmeans(model.host.bio, pairwise ~ timepoint, adjust = "tukey")

# Extract pairwise comparisons
pairwise_comparisons.host.bio <- pairs(pairwise_results.host.bio)

# Filter significant comparisons
significant_comparisons.host.bio <- summary(pairwise_comparisons.host.bio, test = "adjusted", by = NULL, infer = TRUE)

# Extract significant comparisons
significant_contrasts.host.bio <- subset(significant_comparisons.host.bio, p.value < 0.05)

# Make holobiont protein plot
host_bio_cm2 <- master %>%
  filter(!is.na(Host_AFDW.mg.cm2)) %>%
  filter(!is.na(timepoint))%>%
  select(colony_id, timepoint, Host_AFDW.mg.cm2)%>%
  
  ggplot(., aes(x = timepoint, y = Host_AFDW.mg.cm2, fill = timepoint)) +
    geom_boxplot(color="black", fill="white", alpha=0.2) +
    geom_point(pch = 21, size=2, position = position_jitterdodge(0.2)) +
    ylim(0, 3.5)+
  scale_x_discrete(limits = c("JAN","FEB","MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP", "OCT")) +
    xlab("") +
    ylab(expression(bold(paste("Host Biomass (mg cm"^-2, ")"))))+
    theme_classic() + 
    theme(legend.position = "none",
      axis.title=element_text(face="bold", size=60),
      axis.text=element_text(size=40, color="black"), 
      strip.text.x=element_text(face="italic", size=40)); host_bio_cm2



```


Join all plots for each normalization together.  

```{r}
phys_sym_figure<- plot_grid(symbplot_cm2, chla_cm2, chlc_prot_cm2, ncol=3, nrow=1, labels = c('A', 'B', 'C'),label_size = 80,rel_heights= 1, rel_widths = 1, label_y=1, align="h") 

#add x label to plot_GRID
x.grob <- textGrob("TIMEPOINT", 
                   gp=gpar(fontface="bold", col="black", fontsize=50))

#add to plot
phys_sym_figure <- grid.arrange(arrangeGrob(phys_sym_figure, bottom = x.grob))

ggsave(filename="gametogenesis_timeseries_analysis/Figures/Phys_Sym_Figure.pdf", plot=phys_sym_figure, dpi=300, width=49, height=20, units="in")
ggsave(filename="gametogenesis_timeseries_analysis/Figures/Phys_Sym_Figure.png", plot=phys_sym_figure, dpi=300, width=49, height=20, units="in")

phys_host_figure<-plot_grid( host_prot_cm2, holobiont_prot_cm2, ncol=2, nrow=1, labels = c('A', 'B'),label_size = 80,rel_heights= 1, rel_widths = 1, label_y=1, align="h")

#add x label to plot_GRID
x.grob <- textGrob("TIMEPOINT", 
                   gp=gpar(fontface="bold", col="black", fontsize=50))

#add to plot
phys_host_figure <- grid.arrange(arrangeGrob(phys_host_figure, bottom = x.grob))

ggsave(filename="gametogenesis_timeseries_analysis/Figures/Phys_Host_Figure.pdf", plot=phys_host_figure, dpi=300, width=35, height=20, units="in")
ggsave(filename="gametogenesis_timeseries_analysis/Figures/Phys_Host_Figure.png", plot=phys_host_figure, dpi=300, width=35, height=20, units="in")
```

