---
title: "Multivariate analysis of gametogenesis physiological, reproductive, and environmental data"
author: "Ariana S Huffmyer, E5 RoL Team"
edited: "DM Becker"
date: "20240709"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")
# Install devtools if not already installed
if (!require("devtools")) install.packages("devtools")
# Load devtools
library(devtools)
# Install pairwiseAdonis from GitHub
devtools::install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
if (!require("corrplot")) install.packages("corrplot")
if (!require("dplyr")) install.packages("dplyr")

# load packages
library(tidyverse)
library(ggrepel)
library(plyr)
library(corrplot)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(lubridate)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(dplyr)
library(corrr)
library(ggcorrplot)
library(GGally)
library(broom)
library(cowplot)
library(patchwork)
library(pairwiseAdonis)
```

# Load dataframe

Use univariate analyses plots for main figure, comment this out for now if we want dot plots with mean se
Load in master data frames for environmental, physiological, and reproductive data 
```{r}
# master<-read.csv("gametogenesis_timeseries_analysis/Output/master_timeseries.csv")
```

# Aggregate and rename physioloigcal data
```{r}
# # Assuming phys_data is already filtered and renamed correctly
# phys_data <- master %>%
#   dplyr::select(colony_id, timepoint, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Total_Chl_cm2, host_prot_mg.cm2, holobiont_prot_mg.cm2, cells.cm2) %>%
#   filter(Host_AFDW.mg.cm2 > 0) %>%
#   filter(Sym_AFDW.mg.cm2 > 0) %>%
#   dplyr::rename(
#     Host_Biomass = Host_AFDW.mg.cm2,
#     Symbiont_Biomass = Sym_AFDW.mg.cm2,
#     Symbiont_Density = cells.cm2,
#     Host_Protein = host_prot_mg.cm2,
#     Holobiont_Protein = holobiont_prot_mg.cm2,
#     Total_Chlorophyll = Total_Chl_cm2)
# 
# # Ensure timepoint is a factor with appropriate levels
# phys_data$timepoint <- factor(phys_data$timepoint, levels = unique(phys_data$timepoint))
# 
# # Calculate IQR for Total_Chlorophyll due to outliers
# Q1 <- quantile(phys_data$Total_Chlorophyll, 0.25, na.rm = TRUE)
# Q3 <- quantile(phys_data$Total_Chlorophyll, 0.75, na.rm = TRUE)
# IQR <- Q3 - Q1
# 
# # Define bounds for outliers
# lower_bound <- Q1 - 1.5 * IQR
# upper_bound <- Q3 + 1.5 * IQR
# 
# # Remove outliers
# phys_data_cleaned <- phys_data %>%
#   filter(Total_Chlorophyll >= lower_bound & Total_Chlorophyll <= upper_bound)
# 
# # Remove August ACR-426 datapoint from the cleaned data for host protein due to outlier
# phys_data_cleaned <- phys_data_cleaned %>%
#   filter(!(timepoint == "AUGUST" & colony_id == "ACR-426"))
# 
# # Calculate mean, se and SD per timepoint
# summary_data <- do.call(rbind, by(phys_data_cleaned, phys_data_cleaned$timepoint, function(x) {
#   n <- nrow(x)  # Sample size
#   
#   calculate_stats <- function(var) {
#     mean_val <- mean(var, na.rm = TRUE)
#     sd_val <- sd(var, na.rm = TRUE)
#     se_val <- sd_val / sqrt(n)
#     return(c(mean = mean_val, sd = sd_val, se = se_val))}
#   
#   stats_list <- list(
#     Host_Biomass = calculate_stats(x$Host_Biomass),
#     Symbiont_Biomass = calculate_stats(x$Symbiont_Biomass),
#     Total_Chlorophyll = calculate_stats(x$Total_Chlorophyll),
#     Host_Protein = calculate_stats(x$Host_Protein),
#     Holobiont_Protein = calculate_stats(x$Holobiont_Protein),
#     Symbiont_Density = calculate_stats(x$Symbiont_Density))
#   
#   result <- data.frame(timepoint = unique(x$timepoint))
#   for (var in names(stats_list)) {
#     result[[paste0(var, "_mean")]] <- stats_list[[var]]["mean"]
#     result[[paste0(var, "_sd")]] <- stats_list[[var]]["sd"]
#     result[[paste0(var, "_se")]] <- stats_list[[var]]["se"]}
#   
#   return(result)}))
# 
# # Convert row names to a column if needed
# summary_data$timepoint <- rownames(summary_data)
# rownames(summary_data) <- NULL
# 
# # Set the timepoint factor levels in the correct order
# summary_data$timepoint <- factor(summary_data$timepoint, 
#                                   levels = c("JANUARY", "FEBRUARY", "MARCH", 
#                                              "APRIL", "MAY", "JUNE", 
#                                              "JULY", "AUGUST", "SEPTEMBER", 
#                                              "OCTOBER"))
```

# Load in histological data
```{r}
#load histo data
oogen_abundance <- read.csv("gametogenesis_timeseries_analysis/Output/oocyte_abun_data.csv")[-1]
  
sperm_abundance <- read.csv("gametogenesis_timeseries_analysis/Output/sperm_abun_data.csv")[-1]

oogen_size <- read.csv("gametogenesis_timeseries_analysis/Output/oogen_size_data.csv")[-1]
  
sperm_size <- read.csv("gametogenesis_timeseries_analysis/Output/sperm_size_data.csv")[-1]

```

# Load in light and temperature data
```{r}
light.dat <- read.csv("../RAnalysis/output/environmental/mahana_light_data.csv")[, -1]

temp.dat <- read.csv("gametogenesis_timeseries_analysis/Output/mean_temp_data.csv")[, -1]

```

Create downstream plots to use in fully combined figure
Smoothed temperature plot
```{r}
# Convert date.time to POSIXct format
temp.dat$date.time <- as.POSIXct(temp.dat$date.time, format="%Y-%m-%d %H:%M", tz="UTC")

#filter to not show november 
temp.dat <- temp.dat %>%
  filter(month != "Nov")

# Smoothed temperature plot across time
temp_summary_plot <- temp.dat %>%
  ggplot(., aes(x=date.time, y= temp)) + 
  geom_point(colour="gray", size=0.5, alpha=0.05) + 
  geom_smooth(aes(), se=FALSE, method="gam", colour = "black") + 
  ylab("Temperature °C") + 
  xlab("") +
  scale_x_datetime(date_breaks = "1 months", date_labels = "%b %Y")+
  theme_classic() + 
  theme(
    text=element_text(size=14, colour="black"), 
    axis.title=element_text(size=18, colour="black", face="bold"), 
    axis.text.x = element_blank(), # Hide x-axis text labels
    legend.title=element_text(size=18, colour="black", face="bold")) +
      ylim(25, 30);temp_summary_plot

```


Smoothed light plot 
```{r}
# Combine Date and Time columns into a new column named DateTime
light.dat$DateTime <- with(light.dat, ymd_hms(paste(Date, Time)))

light_summary_plot <- light.dat %>%
  ggplot(aes(x = DateTime, y = calibrated)) + 
  geom_point(colour = "darkgray", size = 0.5, alpha = 0.05) + 
  geom_smooth(se = FALSE, colour = "black") + 
  ylab(expression("Light Intensity " * (mu*mol~m^{-2}~s^{-1}))) +
  xlab("") + # This will set the x-axis label but you can remove this line if you want to remove the x-axis title
  scale_x_datetime(date_breaks = "1 months", date_labels = "%b %Y") +
  theme_classic() + 
  theme(text=element_text(size=14, colour="black"), 
    axis.title=element_text(size=18, colour="black", face="bold"), 
    axis.text.x = element_blank(), # Hide x-axis text labels
    legend.title=element_text(size=18, colour="black", face="bold")) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_cartesian(ylim = c(1000, 6000));light_summary_plot

```

Physiological variable plots
Symbiont densities
```{r}
# Symbiont densities plot
sym <- ggplot(summary_data, aes(x = timepoint, y = Symbiont_Density_mean)) +
  geom_point(size = 3, color = "black") +
  geom_errorbar(aes(ymin = Symbiont_Density_mean - Symbiont_Density_se, ymax = Symbiont_Density_mean + Symbiont_Density_se), width = 0.2, color = "black") +
  xlab("") +
  ylab(expression(bold(paste("Symbiont Densities cm"^-2)))) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold", size = 20),
    axis.text.x = element_blank(),         # Remove x-axis text labels
    axis.ticks.x = element_blank(),        # Remove x-axis ticks
    axis.text.y = element_text(size = 20, color = "black"), # Y-axis text labels
    axis.ticks.y = element_line(color = "black"), # Y-axis ticks
    strip.text.x = element_text(face = "italic", size = 20),
    panel.grid.major = element_blank(),    # Remove major gridlines
    panel.grid.minor = element_blank(),    # Remove minor gridlines
    panel.border = element_blank(),        # Remove panel border
    axis.line = element_line(color = "black"))

```

Chlorophyll plot
```{r}
# Plot for Total Chlorophyll
chl <- ggplot(summary_data, aes(x = timepoint, y = Total_Chlorophyll_mean)) +
  geom_point(size = 3, color = "black") +
  geom_errorbar(aes(ymin = Total_Chlorophyll_mean - Total_Chlorophyll_se, ymax = Total_Chlorophyll_mean + Total_Chlorophyll_se), width = 0.2, color = "black") +
  geom_smooth(method = "lm", color = "blue", se = FALSE, linetype = "solid") +  # Adding a linear trendline
  xlab("") +
  ylab(expression(bold(paste("Total Chlorophyll (", mu, "g cm"^-2, ")")))) +  # y-axis label with Greek letter
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold", size = 20),
    axis.text.x = element_blank(),         # Remove x-axis text labels
    axis.ticks.x = element_blank(),        # Remove x-axis ticks
    axis.text.y = element_text(size = 20, color = "black"), # Y-axis text labels
    axis.ticks.y = element_line(color = "black"), # Y-axis ticks
    strip.text.x = element_text(face = "italic", size = 20),
    panel.grid.major = element_blank(),    # Remove major gridlines
    panel.grid.minor = element_blank(),    # Remove minor gridlines
    panel.border = element_blank(),        # Remove panel border
    axis.line = element_line(color = "black"))  # Add axis lines

```

Symbiont biomass
```{r}
# Plot for Symbiont Biomass
sym.bio <- ggplot(summary_data, aes(x = timepoint, y = Symbiont_Biomass_mean)) +
  geom_point(size = 3, color = "black") +
  geom_errorbar(aes(ymin = Symbiont_Biomass_mean - Symbiont_Biomass_se, ymax = Symbiont_Biomass_mean + Symbiont_Biomass_se), width = 0.2, color = "black") +
  geom_smooth(method = "lm", color = "blue", se = FALSE, linetype = "solid") +  # Adding a linear trendline
  xlab("") +
  ylab(expression(bold(paste("Symbiont Biomass (mg cm"^-2, ")")))) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold", size = 20),
    axis.text.x = element_blank(),         # Remove x-axis text labels
    axis.ticks.x = element_blank(),        # Remove x-axis ticks
    axis.text.y = element_text(size = 20, color = "black"), # Y-axis text labels
    axis.ticks.y = element_line(color = "black"), # Y-axis ticks
    strip.text.x = element_text(face = "italic", size = 20),
    panel.grid.major = element_blank(),    # Remove major gridlines
    panel.grid.minor = element_blank(),    # Remove minor gridlines
    panel.border = element_blank(),        # Remove panel border
    axis.line = element_line(color = "black"))  # Add axis lines

```

Host biomass
```{r}
# Plot for Host Biomass
host.bio <- ggplot(summary_data, aes(x = timepoint, y = Host_Biomass_mean)) +
  geom_point(size = 3, color = "black") +
  geom_errorbar(aes(ymin = Host_Biomass_mean - Host_Biomass_se, ymax = Host_Biomass_mean + Host_Biomass_se), width = 0.2, color = "black") +
  geom_smooth(method = "lm", color = "blue", se = FALSE, linetype = "solid") +  # Adding a linear trendline
  xlab("") +
  ylab(expression(bold(paste("Host Biomass (mg cm"^-2, ")")))) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold", size = 20),
    axis.text.x = element_blank(),         # Remove x-axis text labels
    axis.ticks.x = element_blank(),        # Remove x-axis ticks
    axis.text.y = element_text(size = 20, color = "black"), # Y-axis text labels
    axis.ticks.y = element_line(color = "black"), # Y-axis ticks
    strip.text.x = element_text(face = "italic", size = 20),
    panel.grid.major = element_blank(),    # Remove major gridlines
    panel.grid.minor = element_blank(),    # Remove minor gridlines
    panel.border = element_blank(),        # Remove panel border
    axis.line = element_line(color = "black"))  # Add axis lines

```


Host protein
```{r}
# Host protein
host_protein <- ggplot(summary_data, aes(x = timepoint, y = Host_Protein_mean)) +
  geom_point(size = 3, color = "black") +
  geom_errorbar(aes(ymin = Host_Protein_mean - Host_Protein_se, ymax = Host_Protein_mean + Host_Protein_se), width = 0.2, color = "black") +
  geom_smooth(method = "lm", color = "blue", se = FALSE, linetype = "solid") +  # Adding a linear trendline
  xlab("") +
  ylab(expression(bold(paste("Host Protein (mg cm"^-2, ")")))) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold", size = 20),
    axis.text.x = element_blank(),         # Remove x-axis text labels
    axis.ticks.x = element_blank(),        # Remove x-axis ticks
    axis.text.y = element_text(size = 20, color = "black"), # Y-axis text labels
    axis.ticks.y = element_line(color = "black"), # Y-axis ticks
    strip.text.x = element_text(face = "italic", size = 20),
    panel.grid.major = element_blank(),    # Remove major gridlines
    panel.grid.minor = element_blank(),    # Remove minor gridlines
    panel.border = element_blank(),        # Remove panel border
    axis.line = element_line(color = "black"))  # Add axis lines

```


Holobiont protein
```{r}
# Plot for Holobiont Protein
holo_protein <- ggplot(summary_data, aes(x = timepoint, y = Holobiont_Protein_mean)) +
  geom_point(size = 3, color = "black") +
  geom_errorbar(aes(ymin = Holobiont_Protein_mean - Holobiont_Protein_se, ymax = Holobiont_Protein_mean + Holobiont_Protein_se), width = 0.2, color = "black") +
  geom_smooth(method = "lm", color = "blue", se = FALSE, linetype = "solid") +  # Adding a linear trendline
  xlab("") +
  ylab(expression(bold(paste("Holobiont Protein (mg cm"^-2, ")")))) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold", size = 20),
    axis.text.x = element_blank(),         # Remove x-axis text labels
    axis.ticks.x = element_blank(),        # Remove x-axis ticks
    axis.text.y = element_text(size = 20, color = "black"), # Y-axis text labels
    axis.ticks.y = element_line(color = "black"), # Y-axis ticks
    strip.text.x = element_text(face = "italic", size = 20),
    panel.grid.major = element_blank(),    # Remove major gridlines
    panel.grid.minor = element_blank(),    # Remove minor gridlines
    panel.border = element_blank(),        # Remove panel border
    axis.line = element_line(color = "black"))  # Add axis lines

```


```{r}
# Customize environmental and physiological plots to remove x-axis labels and ticks except for the bottom plot
temp_summary_plot <- temp_summary_plot + theme(axis.title.x = element_blank(), 
                                     axis.text.x = element_blank(), 
                                     axis.ticks.x = element_blank(),
                                     axis.title=element_text(face="bold", size=18),
                                     axis.text=element_text(size=18, color="black"))

light_summary_plot <- light_summary_plot + theme(axis.title.x = element_blank(), 
                                   axis.text.x = element_blank(), 
                                   axis.ticks.x = element_blank(),
                                   axis.title=element_text(face="bold", size=18),
                                   axis.text=element_text(size=18, color="black"))

symbplot_cm2 <- symbplot_cm2 + theme(axis.title.x = element_blank(), 
                                     axis.text.x = element_blank(), 
                                     axis.ticks.x = element_blank(),
                                     axis.title=element_text(face="bold", size=18),
                                     axis.text=element_text(size=18, color="black"))

sym_bio_cm2 <- sym_bio_cm2 + theme(axis.title.x = element_blank(), 
                                   axis.text.x = element_blank(), 
                                   axis.ticks.x = element_blank(),
                                     axis.title=element_text(face="bold", size=18),
                                     axis.text=element_text(size=18, color="black"))

tot.chl <- tot.chl + theme(axis.title.x = element_blank(), 
                           axis.text.x = element_blank(), 
                           axis.ticks.x = element_blank(),
                           axis.title=element_text(face="bold", size=18),
                           axis.text=element_text(size=18, color="black"))

host_prot_cm2 <- host_prot_cm2 + theme(axis.title.x = element_blank(), 
                                       axis.text.x = element_blank(), 
                                       axis.ticks.x = element_blank(),
                                       axis.title=element_text(face="bold", size=18),
                                       axis.text=element_text(size=18, color="black"))


holobiont_prot_cm2 <- holobiont_prot_cm2 + theme(axis.title.x = element_blank(), 
                                                 axis.text.x = element_blank(), 
                                                 axis.ticks.x = element_blank(),
                                                 axis.title=element_text(face="bold", size=18),
                                                 axis.text=element_text(size=18, color="black"))

host_bio_cm2 <- host_bio_cm2 + theme(axis.title=element_text(face="bold", size=18),
                                                 axis.text=element_text(size=18, color="black"))

# Leave the x-axis labels and ticks for the last plot (host_bio_cm2)
# No changes needed for host_bio_cm2 as it will display the x-axis

# Combine the physiological and environmmental plots (with x-axis labels only on the bottom plot)
combined_plot <- (temp_summary_plot / light_summary_plot / symbplot_cm2 / sym_bio_cm2 / tot.chl / host_prot_cm2 / holobiont_prot_cm2 / host_bio_cm2) +
  plot_layout(ncol = 1, guides = 'collect')

# Save combined plot
ggsave(filename="gametogenesis_timeseries_analysis/Figures/timeseries_stacked_plot.png", plot=combined_plot, dpi=300, width=10, height=20, units="in")
```




