---
title: "Multivariate analysis of gametogenesis physiological, reproductive, and environmental data"
author: "Ariana S Huffmyer, E5 RoL Team"
edited: "DM Becker"
date: "20240709"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 6
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

# Set Up    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
## install packages if you dont already have them
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("RColorBrewer")) install.packages("RColorBrewer")
if (!require("lme4")) install.packages("lme4")
if (!require("lmerTest")) install.packages("lmerTest")
if (!require("car")) install.packages("car")
if (!require("effects")) install.packages("effects")
if (!require("ggfortify")) install.packages("ggfortify")
if (!require("cowplot")) install.packages("cowplot")
if (!require("vegan")) install.packages("vegan")
if (!require("corrr")) install.packages("corrr")
if (!require("ggcorrplot")) install.packages("ggcorrplot")
if (!require("GGally")) install.packages("GGally")
if (!require("broom")) install.packages("broom")
if (!require("cowplot")) install.packages("cowplot")
# Install devtools if not already installed
if (!require("devtools")) install.packages("devtools")
# Load devtools
library(devtools)
# Install pairwiseAdonis from GitHub
devtools::install_github("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
if (!require("corrplot")) install.packages("corrplot")
if (!require("dplyr")) install.packages("dplyr")
if (!require("mctest")) install.packages("mctest")
if (!require("ordinal")) install.packages("ordinal")

# load packages
library(tidyverse)
library(lmtest)
library(ordinal)
library(ggrepel)
library(mctest)
library(MASS)
library(VGAM)
library(brant)
library(plyr)
library(corrplot)
library(ggplot2)
library(RColorBrewer)
library(lme4)
library(lmerTest)
library(car)
library(lubridate)
library(effects)
library(ggfortify)
library(cowplot)
library(vegan)
library(dplyr)
library(corrr)
library(ggcorrplot)
library(GGally)
library(broom)
library(cowplot)
library(patchwork)
library(pairwiseAdonis)
```

# Load dataframe

Use univariate analyses plots for main figure, comment this out for now if we want dot plots with mean se
Load in master data frames for environmental, physiological, and reproductive data 
```{r}
# master<-read.csv("gametogenesis_timeseries_analysis/Output/master_timeseries.csv")
```

# Aggregate and rename physioloigcal data
```{r}
# # Assuming phys_data is already filtered and renamed correctly
# phys_data <- master %>%
#   dplyr::select(colony_id, timepoint, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Total_Chl_cm2, host_prot_mg.cm2, holobiont_prot_mg.cm2, cells.cm2) %>%
#   filter(Host_AFDW.mg.cm2 > 0) %>%
#   filter(Sym_AFDW.mg.cm2 > 0) %>%
#   dplyr::rename(
#     Host_Biomass = Host_AFDW.mg.cm2,
#     Symbiont_Biomass = Sym_AFDW.mg.cm2,
#     Symbiont_Density = cells.cm2,
#     Host_Protein = host_prot_mg.cm2,
#     Holobiont_Protein = holobiont_prot_mg.cm2,
#     Total_Chlorophyll = Total_Chl_cm2)
# 
# # Ensure timepoint is a factor with appropriate levels
# phys_data$timepoint <- factor(phys_data$timepoint, levels = unique(phys_data$timepoint))
# 
# # Calculate IQR for Total_Chlorophyll due to outliers
# Q1 <- quantile(phys_data$Total_Chlorophyll, 0.25, na.rm = TRUE)
# Q3 <- quantile(phys_data$Total_Chlorophyll, 0.75, na.rm = TRUE)
# IQR <- Q3 - Q1
# 
# # Define bounds for outliers
# lower_bound <- Q1 - 1.5 * IQR
# upper_bound <- Q3 + 1.5 * IQR
# 
# # Remove outliers
# phys_data_cleaned <- phys_data %>%
#   filter(Total_Chlorophyll >= lower_bound & Total_Chlorophyll <= upper_bound)
# 
# # Remove August ACR-426 datapoint from the cleaned data for host protein due to outlier
# phys_data_cleaned <- phys_data_cleaned %>%
#   filter(!(timepoint == "AUGUST" & colony_id == "ACR-426"))
# 
# # Calculate mean, se and SD per timepoint
# summary_data <- do.call(rbind, by(phys_data_cleaned, phys_data_cleaned$timepoint, function(x) {
#   n <- nrow(x)  # Sample size
#   
#   calculate_stats <- function(var) {
#     mean_val <- mean(var, na.rm = TRUE)
#     sd_val <- sd(var, na.rm = TRUE)
#     se_val <- sd_val / sqrt(n)
#     return(c(mean = mean_val, sd = sd_val, se = se_val))}
#   
#   stats_list <- list(
#     Host_Biomass = calculate_stats(x$Host_Biomass),
#     Symbiont_Biomass = calculate_stats(x$Symbiont_Biomass),
#     Total_Chlorophyll = calculate_stats(x$Total_Chlorophyll),
#     Host_Protein = calculate_stats(x$Host_Protein),
#     Holobiont_Protein = calculate_stats(x$Holobiont_Protein),
#     Symbiont_Density = calculate_stats(x$Symbiont_Density))
#   
#   result <- data.frame(timepoint = unique(x$timepoint))
#   for (var in names(stats_list)) {
#     result[[paste0(var, "_mean")]] <- stats_list[[var]]["mean"]
#     result[[paste0(var, "_sd")]] <- stats_list[[var]]["sd"]
#     result[[paste0(var, "_se")]] <- stats_list[[var]]["se"]}
#   
#   return(result)}))
# 
# # Convert row names to a column if needed
# summary_data$timepoint <- rownames(summary_data)
# rownames(summary_data) <- NULL
# 
# # Set the timepoint factor levels in the correct order
# summary_data$timepoint <- factor(summary_data$timepoint, 
#                                   levels = c("JANUARY", "FEBRUARY", "MARCH", 
#                                              "APRIL", "MAY", "JUNE", 
#                                              "JULY", "AUGUST", "SEPTEMBER", 
#                                              "OCTOBER"))
```

# Load in histological data
```{r}
#load histo data
oogen_abundance <- read.csv("gametogenesis_timeseries_analysis/Output/oocyte_abun_data.csv")[-1]
  
sperm_abundance <- read.csv("gametogenesis_timeseries_analysis/Output/sperm_abun_data.csv")[-1]

oogen_size <- read.csv("gametogenesis_timeseries_analysis/Output/oogen_size_data.csv")[-1]
  
sperm_size <- read.csv("gametogenesis_timeseries_analysis/Output/sperm_size_data.csv")[-1]

```

# Load in light and temperature data
```{r}
light.dat <- read.csv("../RAnalysis/output/environmental/mahana_light_data.csv")[, -1]

temp.dat <- read.csv("gametogenesis_timeseries_analysis/Output/mean_temp_data.csv")[, -1]

```

# Load in physiological data
```{r}
master <- read.csv("gametogenesis_timeseries_analysis/Output/master_timeseries.csv")

# read in metadata to add patch to master timeseries
metadata <- read.csv("gametogenesis_timeseries_analysis/master_metadata.csv")

# add ACR- before each tag number
metadata <- metadata %>%
  mutate(colony_id = paste0("ACR-", colony_id))

#commbine master and metadata
master <- left_join(master, metadata)
```

<!-- Create downstream plots to use in fully combined figure -->
<!-- Smoothed temperature plot -->
<!-- ```{r} -->
<!-- # Convert date.time to POSIXct format -->
<!-- temp.dat$date.time <- as.POSIXct(temp.dat$date.time, format="%Y-%m-%d %H:%M", tz="UTC") -->

<!-- #filter to not show november  -->
<!-- temp.dat <- temp.dat %>% -->
<!--   filter(month != "Nov") -->

<!-- # Smoothed temperature plot across time -->
<!-- temp_summary_plot <- temp.dat %>% -->
<!--   ggplot(., aes(x=date.time, y= temp)) +  -->
<!--   geom_point(colour="gray", size=0.5, alpha=0.05) +  -->
<!--   geom_smooth(aes(), se=FALSE, method="gam", colour = "black") +  -->
<!--   ylab("Temperature °C") +  -->
<!--   xlab("") + -->
<!--   scale_x_datetime(date_breaks = "1 months", date_labels = "%b %Y")+ -->
<!--   theme_classic() +  -->
<!--   theme( -->
<!--     text=element_text(size=14, colour="black"),  -->
<!--     axis.title=element_text(size=18, colour="black", face="bold"),  -->
<!--     axis.text.x = element_blank(), # Hide x-axis text labels -->
<!--     legend.title=element_text(size=18, colour="black", face="bold")) + -->
<!--       ylim(25, 30);temp_summary_plot -->

<!-- ``` -->


<!-- Smoothed light plot  -->
<!-- ```{r} -->
<!-- # Combine Date and Time columns into a new column named DateTime -->
<!-- light.dat$DateTime <- with(light.dat, ymd_hms(paste(Date, Time))) -->

<!-- light_summary_plot <- light.dat %>% -->
<!--   ggplot(aes(x = DateTime, y = calibrated)) +  -->
<!--   geom_point(colour = "darkgray", size = 0.5, alpha = 0.05) +  -->
<!--   geom_smooth(se = FALSE, colour = "black") +  -->
<!--   ylab(expression("Light Intensity " * (mu*mol~m^{-2}~s^{-1}))) + -->
<!--   xlab("") + # This will set the x-axis label but you can remove this line if you want to remove the x-axis title -->
<!--   scale_x_datetime(date_breaks = "1 months", date_labels = "%b %Y") + -->
<!--   theme_classic() +  -->
<!--   theme(text=element_text(size=14, colour="black"),  -->
<!--     axis.title=element_text(size=18, colour="black", face="bold"),  -->
<!--     axis.text.x = element_blank(), # Hide x-axis text labels -->
<!--     legend.title=element_text(size=18, colour="black", face="bold")) + -->
<!--   scale_y_continuous(expand = c(0, 0)) + -->
<!--   coord_cartesian(ylim = c(1000, 6000));light_summary_plot -->

<!-- ``` -->

<!-- Physiological variable plots -->
<!-- Symbiont densities -->
<!-- ```{r} -->
<!-- # Symbiont densities plot -->
<!-- sym <- ggplot(summary_data, aes(x = timepoint, y = Symbiont_Density_mean)) + -->
<!--   geom_point(size = 3, color = "black") + -->
<!--   geom_errorbar(aes(ymin = Symbiont_Density_mean - Symbiont_Density_se, ymax = Symbiont_Density_mean + Symbiont_Density_se), width = 0.2, color = "black") + -->
<!--   xlab("") + -->
<!--   ylab(expression(bold(paste("Symbiont Densities cm"^-2)))) + -->
<!--   theme_minimal() + -->
<!--   theme( -->
<!--     legend.position = "none", -->
<!--     axis.title = element_text(face = "bold", size = 20), -->
<!--     axis.text.x = element_blank(),         # Remove x-axis text labels -->
<!--     axis.ticks.x = element_blank(),        # Remove x-axis ticks -->
<!--     axis.text.y = element_text(size = 20, color = "black"), # Y-axis text labels -->
<!--     axis.ticks.y = element_line(color = "black"), # Y-axis ticks -->
<!--     strip.text.x = element_text(face = "italic", size = 20), -->
<!--     panel.grid.major = element_blank(),    # Remove major gridlines -->
<!--     panel.grid.minor = element_blank(),    # Remove minor gridlines -->
<!--     panel.border = element_blank(),        # Remove panel border -->
<!--     axis.line = element_line(color = "black")) -->

<!-- ``` -->

<!-- Chlorophyll plot -->
<!-- ```{r} -->
<!-- # Plot for Total Chlorophyll -->
<!-- chl <- ggplot(summary_data, aes(x = timepoint, y = Total_Chlorophyll_mean)) + -->
<!--   geom_point(size = 3, color = "black") + -->
<!--   geom_errorbar(aes(ymin = Total_Chlorophyll_mean - Total_Chlorophyll_se, ymax = Total_Chlorophyll_mean + Total_Chlorophyll_se), width = 0.2, color = "black") + -->
<!--   geom_smooth(method = "lm", color = "blue", se = FALSE, linetype = "solid") +  # Adding a linear trendline -->
<!--   xlab("") + -->
<!--   ylab(expression(bold(paste("Total Chlorophyll (", mu, "g cm"^-2, ")")))) +  # y-axis label with Greek letter -->
<!--   theme_minimal() + -->
<!--   theme( -->
<!--     legend.position = "none", -->
<!--     axis.title = element_text(face = "bold", size = 20), -->
<!--     axis.text.x = element_blank(),         # Remove x-axis text labels -->
<!--     axis.ticks.x = element_blank(),        # Remove x-axis ticks -->
<!--     axis.text.y = element_text(size = 20, color = "black"), # Y-axis text labels -->
<!--     axis.ticks.y = element_line(color = "black"), # Y-axis ticks -->
<!--     strip.text.x = element_text(face = "italic", size = 20), -->
<!--     panel.grid.major = element_blank(),    # Remove major gridlines -->
<!--     panel.grid.minor = element_blank(),    # Remove minor gridlines -->
<!--     panel.border = element_blank(),        # Remove panel border -->
<!--     axis.line = element_line(color = "black"))  # Add axis lines -->

<!-- ``` -->

<!-- Symbiont biomass -->
<!-- ```{r} -->
<!-- # Plot for Symbiont Biomass -->
<!-- sym.bio <- ggplot(summary_data, aes(x = timepoint, y = Symbiont_Biomass_mean)) + -->
<!--   geom_point(size = 3, color = "black") + -->
<!--   geom_errorbar(aes(ymin = Symbiont_Biomass_mean - Symbiont_Biomass_se, ymax = Symbiont_Biomass_mean + Symbiont_Biomass_se), width = 0.2, color = "black") + -->
<!--   geom_smooth(method = "lm", color = "blue", se = FALSE, linetype = "solid") +  # Adding a linear trendline -->
<!--   xlab("") + -->
<!--   ylab(expression(bold(paste("Symbiont Biomass (mg cm"^-2, ")")))) + -->
<!--   theme_minimal() + -->
<!--   theme( -->
<!--     legend.position = "none", -->
<!--     axis.title = element_text(face = "bold", size = 20), -->
<!--     axis.text.x = element_blank(),         # Remove x-axis text labels -->
<!--     axis.ticks.x = element_blank(),        # Remove x-axis ticks -->
<!--     axis.text.y = element_text(size = 20, color = "black"), # Y-axis text labels -->
<!--     axis.ticks.y = element_line(color = "black"), # Y-axis ticks -->
<!--     strip.text.x = element_text(face = "italic", size = 20), -->
<!--     panel.grid.major = element_blank(),    # Remove major gridlines -->
<!--     panel.grid.minor = element_blank(),    # Remove minor gridlines -->
<!--     panel.border = element_blank(),        # Remove panel border -->
<!--     axis.line = element_line(color = "black"))  # Add axis lines -->

<!-- ``` -->

<!-- Host biomass -->
<!-- ```{r} -->
<!-- # Plot for Host Biomass -->
<!-- host.bio <- ggplot(summary_data, aes(x = timepoint, y = Host_Biomass_mean)) + -->
<!--   geom_point(size = 3, color = "black") + -->
<!--   geom_errorbar(aes(ymin = Host_Biomass_mean - Host_Biomass_se, ymax = Host_Biomass_mean + Host_Biomass_se), width = 0.2, color = "black") + -->
<!--   geom_smooth(method = "lm", color = "blue", se = FALSE, linetype = "solid") +  # Adding a linear trendline -->
<!--   xlab("") + -->
<!--   ylab(expression(bold(paste("Host Biomass (mg cm"^-2, ")")))) + -->
<!--   theme_minimal() + -->
<!--   theme( -->
<!--     legend.position = "none", -->
<!--     axis.title = element_text(face = "bold", size = 20), -->
<!--     axis.text.x = element_blank(),         # Remove x-axis text labels -->
<!--     axis.ticks.x = element_blank(),        # Remove x-axis ticks -->
<!--     axis.text.y = element_text(size = 20, color = "black"), # Y-axis text labels -->
<!--     axis.ticks.y = element_line(color = "black"), # Y-axis ticks -->
<!--     strip.text.x = element_text(face = "italic", size = 20), -->
<!--     panel.grid.major = element_blank(),    # Remove major gridlines -->
<!--     panel.grid.minor = element_blank(),    # Remove minor gridlines -->
<!--     panel.border = element_blank(),        # Remove panel border -->
<!--     axis.line = element_line(color = "black"))  # Add axis lines -->

<!-- ``` -->


<!-- Host protein -->
<!-- ```{r} -->
<!-- # Host protein -->
<!-- host_protein <- ggplot(summary_data, aes(x = timepoint, y = Host_Protein_mean)) + -->
<!--   geom_point(size = 3, color = "black") + -->
<!--   geom_errorbar(aes(ymin = Host_Protein_mean - Host_Protein_se, ymax = Host_Protein_mean + Host_Protein_se), width = 0.2, color = "black") + -->
<!--   geom_smooth(method = "lm", color = "blue", se = FALSE, linetype = "solid") +  # Adding a linear trendline -->
<!--   xlab("") + -->
<!--   ylab(expression(bold(paste("Host Protein (mg cm"^-2, ")")))) + -->
<!--   theme_minimal() + -->
<!--   theme( -->
<!--     legend.position = "none", -->
<!--     axis.title = element_text(face = "bold", size = 20), -->
<!--     axis.text.x = element_blank(),         # Remove x-axis text labels -->
<!--     axis.ticks.x = element_blank(),        # Remove x-axis ticks -->
<!--     axis.text.y = element_text(size = 20, color = "black"), # Y-axis text labels -->
<!--     axis.ticks.y = element_line(color = "black"), # Y-axis ticks -->
<!--     strip.text.x = element_text(face = "italic", size = 20), -->
<!--     panel.grid.major = element_blank(),    # Remove major gridlines -->
<!--     panel.grid.minor = element_blank(),    # Remove minor gridlines -->
<!--     panel.border = element_blank(),        # Remove panel border -->
<!--     axis.line = element_line(color = "black"))  # Add axis lines -->

<!-- ``` -->


<!-- Holobiont protein -->
<!-- ```{r} -->
<!-- # Plot for Holobiont Protein -->
<!-- holo_protein <- ggplot(summary_data, aes(x = timepoint, y = Holobiont_Protein_mean)) + -->
<!--   geom_point(size = 3, color = "black") + -->
<!--   geom_errorbar(aes(ymin = Holobiont_Protein_mean - Holobiont_Protein_se, ymax = Holobiont_Protein_mean + Holobiont_Protein_se), width = 0.2, color = "black") + -->
<!--   geom_smooth(method = "lm", color = "blue", se = FALSE, linetype = "solid") +  # Adding a linear trendline -->
<!--   xlab("") + -->
<!--   ylab(expression(bold(paste("Holobiont Protein (mg cm"^-2, ")")))) + -->
<!--   theme_minimal() + -->
<!--   theme( -->
<!--     legend.position = "none", -->
<!--     axis.title = element_text(face = "bold", size = 20), -->
<!--     axis.text.x = element_blank(),         # Remove x-axis text labels -->
<!--     axis.ticks.x = element_blank(),        # Remove x-axis ticks -->
<!--     axis.text.y = element_text(size = 20, color = "black"), # Y-axis text labels -->
<!--     axis.ticks.y = element_line(color = "black"), # Y-axis ticks -->
<!--     strip.text.x = element_text(face = "italic", size = 20), -->
<!--     panel.grid.major = element_blank(),    # Remove major gridlines -->
<!--     panel.grid.minor = element_blank(),    # Remove minor gridlines -->
<!--     panel.border = element_blank(),        # Remove panel border -->
<!--     axis.line = element_line(color = "black"))  # Add axis lines -->

<!-- ``` -->


<!-- ```{r} -->
<!-- # Customize environmental and physiological plots to remove x-axis labels and ticks except for the bottom plot -->
<!-- temp_summary_plot <- temp_summary_plot + theme(axis.title.x = element_blank(),  -->
<!--                                      axis.text.x = element_blank(),  -->
<!--                                      axis.ticks.x = element_blank(), -->
<!--                                      axis.title=element_text(face="bold", size=18), -->
<!--                                      axis.text=element_text(size=18, color="black")) -->

<!-- light_summary_plot <- light_summary_plot + theme(axis.title.x = element_blank(),  -->
<!--                                    axis.text.x = element_blank(),  -->
<!--                                    axis.ticks.x = element_blank(), -->
<!--                                    axis.title=element_text(face="bold", size=18), -->
<!--                                    axis.text=element_text(size=18, color="black")) -->

<!-- symbplot_cm2 <- symbplot_cm2 + theme(axis.title.x = element_blank(),  -->
<!--                                      axis.text.x = element_blank(),  -->
<!--                                      axis.ticks.x = element_blank(), -->
<!--                                      axis.title=element_text(face="bold", size=18), -->
<!--                                      axis.text=element_text(size=18, color="black")) -->

<!-- sym_bio_cm2 <- sym_bio_cm2 + theme(axis.title.x = element_blank(),  -->
<!--                                    axis.text.x = element_blank(),  -->
<!--                                    axis.ticks.x = element_blank(), -->
<!--                                      axis.title=element_text(face="bold", size=18), -->
<!--                                      axis.text=element_text(size=18, color="black")) -->

<!-- tot.chl <- tot.chl + theme(axis.title.x = element_blank(),  -->
<!--                            axis.text.x = element_blank(),  -->
<!--                            axis.ticks.x = element_blank(), -->
<!--                            axis.title=element_text(face="bold", size=18), -->
<!--                            axis.text=element_text(size=18, color="black")) -->

<!-- host_prot_cm2 <- host_prot_cm2 + theme(axis.title.x = element_blank(),  -->
<!--                                        axis.text.x = element_blank(),  -->
<!--                                        axis.ticks.x = element_blank(), -->
<!--                                        axis.title=element_text(face="bold", size=18), -->
<!--                                        axis.text=element_text(size=18, color="black")) -->


<!-- holobiont_prot_cm2 <- holobiont_prot_cm2 + theme(axis.title.x = element_blank(),  -->
<!--                                                  axis.text.x = element_blank(),  -->
<!--                                                  axis.ticks.x = element_blank(), -->
<!--                                                  axis.title=element_text(face="bold", size=18), -->
<!--                                                  axis.text=element_text(size=18, color="black")) -->

<!-- host_bio_cm2 <- host_bio_cm2 + theme(axis.title=element_text(face="bold", size=18), -->
<!--                                                  axis.text=element_text(size=18, color="black")) -->

<!-- # Leave the x-axis labels and ticks for the last plot (host_bio_cm2) -->
<!-- # No changes needed for host_bio_cm2 as it will display the x-axis -->

<!-- # Combine the physiological and environmmental plots (with x-axis labels only on the bottom plot) -->
<!-- combined_plot <- (temp_summary_plot / light_summary_plot / symbplot_cm2 / sym_bio_cm2 / tot.chl / host_prot_cm2 / holobiont_prot_cm2 / host_bio_cm2) + -->
<!--   plot_layout(ncol = 1, guides = 'collect') -->

<!-- # Save combined plot -->
<!-- ggsave(filename="gametogenesis_timeseries_analysis/Figures/timeseries_stacked_plot.png", plot=combined_plot, dpi=300, width=10, height=20, units="in") -->
<!-- ``` -->


<!-- Run models to look at individual impacts of physio on frequency in stage -->
<!-- ```{r} -->
<!-- # Define the custom function -->
<!-- to_title_case <- function(x) { -->
<!--   gsub("(^|[[:space:]])([[:alpha:]])", "\\1\\U\\2", tolower(x), perl = TRUE) -->
<!-- } -->

<!-- # make timepoints numbered -->
<!-- master_select <- master %>% -->
<!--   mutate(timepoint = to_title_case(timepoint)) %>% -->
<!--   mutate(Month_Number = case_when( -->
<!--     timepoint == "January" ~ 1, -->
<!--     timepoint == "February" ~ 2, -->
<!--     timepoint == "March" ~ 3, -->
<!--     timepoint == "April" ~ 4, -->
<!--     timepoint == "May" ~ 5, -->
<!--     timepoint == "June" ~ 6, -->
<!--     timepoint == "July" ~ 7, -->
<!--     timepoint == "August" ~ 8, -->
<!--     timepoint == "September" ~ 9, -->
<!--     timepoint == "October" ~ 10, -->
<!--     TRUE ~ NA_integer_))  %>%  -->
<!--   dplyr::select(colony_id, Month_Number, timepoint, Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Total_Chl_cm2, host_prot_mg.cm2, holobiont_prot_mg.cm2, cells.cm2) -->

<!-- # get the means per timepoint for all these physiological variables -->
<!-- master_means <- master_select %>% -->
<!--   group_by(timepoint, Month_Number) %>% -->
<!--   dplyr::summarise(across(c(Host_AFDW.mg.cm2, Sym_AFDW.mg.cm2, Total_Chl_cm2,  -->
<!--                      host_prot_mg.cm2, holobiont_prot_mg.cm2, cells.cm2),  -->
<!--                    list(mean = ~mean(., na.rm = TRUE), -->
<!--                         se = ~sd(., na.rm = TRUE) / sqrt(n()))), -->
<!--             .groups = "drop") %>% -->
<!--   arrange(Month_Number) -->


<!-- all.dat <- left_join(master_means, oogen_abundance) -->

<!-- ``` -->


Calculate the condition indices and variance-decomposition proportions for each physiological (explanatory) variable for multicollinearity analysis
```{r}
# Create the subset for the specified variables
subset_dataframe <- master %>%
  dplyr::select(Host_AFDW.mg.cm2, 
         Sym_AFDW.mg.cm2, 
         Total_Chl_cm2, 
         host_prot_mg.cm2, 
         cells.cm2)

# Replace Total_Chl_cm2 values with NA for June due to higher abs values not explained by other biological response variables, could be an artifact in the lab processing
subset_dataframe <- subset_dataframe %>%
  mutate(Total_Chl_cm2 = ifelse(master$timepoint == "JUNE", NA, Total_Chl_cm2))

# Calculate the correlation matrix
correlation_matrix <- cor(subset_dataframe, use = "pairwise.complete.obs")

# Print the correlation matrix
print(correlation_matrix)

# Fit a linear model with all variables
model <- lm(Host_AFDW.mg.cm2 ~ ., data = subset_dataframe)

# Calculate eigenvalues, condition indices, and variance decomposition proportions
multicollinearity_diagnostics <- eigprop(model)

# Print the results
print(multicollinearity_diagnostics)


```
Call:
eigprop(mod = model)

  Eigenvalues     CI (Intercept) Sym_AFDW.mg.cm2 Total_Chl_cm2 host_prot_mg.cm2 cells.cm2
1      3.9585 1.0000      0.0057          0.0164        0.0066           0.0091    0.0090
2      0.7251 2.3365      0.0017          0.9526        0.0056           0.0032    0.0060
3      0.1521 5.1020      0.0041          0.0074        0.0111           0.5623    0.4658
4      0.0936 6.5028      0.0663          0.0024        0.5639           0.2953    0.5092
5      0.0708 7.4795      0.9222          0.0211        0.4128           0.1301    0.0101

===============================
Row 4==> Total_Chl_cm2, proportion 0.563930 >= 0.50 
Row 4==> cells.cm2, proportion 0.509215 >= 0.50 

Condition indices are all < 30 so they are not significant multicollinearlity issues. For variance-decomposition proportions, row 4 for symbiont cells and total chlorophyll can be considered multicollinear due to two variables > 0.50, which makes sense biologically and they are also 41% correlated from the correlation matrix.

Make combined sym and chl variable
```{r}
# add back in timepoint
subset_df <- master %>%
  dplyr::select(Host_AFDW.mg.cm2, 
         Sym_AFDW.mg.cm2, 
         Total_Chl_cm2, 
         host_prot_mg.cm2, 
         cells.cm2, timepoint)

# Assuming your dataframe is called 'df', scale variables
subset_df$Chl_cells_combined <- scale(subset_df$Total_Chl_cm2) + scale(subset_df$cells.cm2)
```


Run ordinal logistic regression for oogenesis and spermatogeneis stages
```{r}
# Assume you have two dataframes: oogenesis_data and spermatogenesis_data
# Adding a new column to identify the process
oogenesis_data <- oogen_abundance %>%
  mutate(Process = "Oogenesis")

spermatogenesis_data <- sperm_abundance %>%
  mutate(Process = "Spermatogenesis")

# Combine the two datasets
combined_data <- bind_rows(oogenesis_data, spermatogenesis_data)

# Make Stage is an ordered factor
combined_data$Stage <- factor(combined_data$Stage, ordered = TRUE, 
                   levels = c("Stage 0", "Stage I", "Stage II", "Stage III", "Stage IV", "Stage V"))

# Convert timepoint to factor for both dfs
combined_data$timepoint <- factor(combined_data$timepoint, 
                       levels = c("January", "February", "March", "April", "May", "June", 
                                  "July", "August", "September", "October", "November", "December"))

#make phys timepoints undercase
subset_df$timepoint <- sapply(subset_df$timepoint, function(x) {
  paste0(toupper(substr(x, 1, 1)), tolower(substr(x, 2, nchar(x))))
})

# Convert timepoint to a factor
subset_df$timepoint <- factor(subset_df$timepoint, 
                       levels = c("January", "February", "March", "April", "May", "June", 
                                  "July", "August", "September", "October", "November", "December"))

# Rename the chl_cells column
subset_df <- subset_df %>%
  rename_with(~ "chl_cells_combined", matches("Chl_cells_combined"))

# Create a weight variable based on frequency
combined_data$weight <- combined_data$freq * 100  # Multiply by 100 to convert to whole numbers

# First, let's check the unique values in the timepoint column
print(unique(subset_df$timepoint))

# Make sure timepoint is a factor with all levels
all_months <- c("January", "February", "March", "April", "May", "June", 
                "July", "August", "September", "October")
# make factor
subset_df$timepoint <- factor(subset_df$timepoint, levels = all_months)

# Merge phys timeseries and repo 
merged_data <- merge(subset_df, combined_data, by = "timepoint", all = TRUE)

# Rename chl_cells column
names(merged_data)[names(merged_data) == "chl_cells_combined[, 1]"] <- "chl_cells_combined"

# Make Stage as an ordered factor
merged_data$Stage <- factor(merged_data$Stage, ordered = TRUE, 
                            levels = c("Stage 0", "Stage I", "Stage II", "Stage III", "Stage IV", "Stage V"))


# Scale all phys variables
merged_data$Host_AFDW.mg.cm2 <- scale(merged_data$Host_AFDW.mg.cm2)
merged_data$Sym_AFDW.mg.cm2 <- scale(merged_data$Sym_AFDW.mg.cm2)
merged_data$host_prot_mg.cm2 <- scale(merged_data$host_prot_mg.cm2)

# Run ordinal logistic regression with sperm and oogen together
model <- polr(Stage ~ Host_AFDW.mg.cm2 + Sym_AFDW.mg.cm2 + host_prot_mg.cm2 + chl_cells_combined + Process, 
              data = merged_data, Hess = TRUE, weights = freq)

summary(model)

# Calculate p-values
ctable <- coef(summary(model))
p <- pnorm(abs(ctable[, "t value"]), lower.tail = FALSE) * 2
ctable <- cbind(ctable, "p value" = p)
print(ctable)

# Test proportional odds assumption
brant(model)

# The Brant test results showed that the ProcessSpermatogenesis variable significantly violated the proportional odds assumption, indicating that its effect varies across different stages of reproduction.

# Run ordinal logistic regressions separately for each gametogenic process
oogenesis_data <- subset(merged_data, Process == "Oogenesis")
spermatogenesis_data <- subset(merged_data, Process == "Spermatogenesis")

# Run oogenesis model
oogenesis_model <- polr(Stage ~ Host_AFDW.mg.cm2 + Sym_AFDW.mg.cm2 + 
                        host_prot_mg.cm2 + chl_cells_combined, 
                        data = oogenesis_data, Hess = TRUE, weights = weight)


summary(oogenesis_model)

# testing parallel regression assumption using Brant's test
brant(oogenesis_model)

# Run spermatogenesis model
spermatogenesis_model <- polr(Stage ~ Host_AFDW.mg.cm2 + Sym_AFDW.mg.cm2 + 
                              host_prot_mg.cm2 + chl_cells_combined, 
                              data = spermatogenesis_data, Hess = TRUE, weights = weight)

summary(spermatogenesis_model)

# testing parallel regression assumption using Brant's test
brant(spermatogenesis_model)


```

Chl_cells and host protein do not meet assumptions of the Brant's test and show variable relationships with stages, try a partial proportional odds model

For oogenesis
```{r}
# Fit the partial proportional odds model
oogenesis_ppom <- vglm(Stage ~ Host_AFDW.mg.cm2 + Sym_AFDW.mg.cm2 + 
                       host_prot_mg.cm2 + chl_cells_combined,
                       family = cumulative(parallel = TRUE ~ Host_AFDW.mg.cm2 + Sym_AFDW.mg.cm2 - 1),
                       data = oogenesis_data)

# Summary of the model
summary(oogenesis_ppom)


# For oogenesis
AIC(oogenesis_ppom)
AIC(oogenesis_model)  # Your original polr model
```

For spermatogenesis
```{r}

```



