---
title: "environmental_processing"
author: "daniellembecker"
date: "2024-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Processing all environmental data collected from Mahana site in Moorea, French Polynesia (continuous light and temperature data) from ~January till November 2022

##Install packages
# load packages

```{r}
library("segmented")
library("devtools")
library("plotrix")
library("gridExtra")
# Install devtools if not already installed
if (!require("devtools")) install.packages("devtools")
# Load devtools
library(devtools)
# Install LoLinR from GitHub
install_github('colin-olito/LoLinR')
library("LoLinR")
library("lubridate")
library('tidyverse')
library('lubridate')
library('scales')
library('patchwork')

```

##Processing temperature data for from HOBO temperature loggers out at Mahana from January to November

#load temp data

```{r}
mahana.temp.hobo.06 <- read.csv("RAnalysis/data/environmental/HOBO_temperature/20220415_Mahana_6.csv")
mahana.temp.hobo.09 <- read.csv("RAnalysis/data/environmental/HOBO_temperature/20220415_Mahana_9.csv")
mahana.temp.hobo.07 <- read.csv("RAnalysis/data/environmental/HOBO_temperature/20220415_Mahana_7.csv")
mahana.temp.hobo.10 <- read.csv("RAnalysis/data/environmental/HOBO_temperature/20221114_Mahana_10.csv")
mahana.temp.hobo.11 <- read.csv("RAnalysis/data/environmental/HOBO_temperature/20221114_Mahana_11.csv")
mahana.temp.hobo.12 <- read.csv("RAnalysis/data/environmental/HOBO_temperature/20221114_Mahana_12.csv")
mahana.temp.hobo.13 <- read.csv("RAnalysis/data/environmental/HOBO_temperature/20221114_Mahana_13.csv")
mahana.temp.hobo.08 <- read.csv("RAnalysis/data/environmental/HOBO_temperature/20221114_Mahana_8.csv")
mahana.temp.hobo.05 <- read.csv("RAnalysis/data/environmental/HOBO_temperature/20221114_Mahana_5.csv")
mahana.temp.hobo.04 <- read.csv("RAnalysis/data/environmental/HOBO_temperature/20221114_Mahana_4.csv")
mahana.temp.hobo.03 <- read.csv("RAnalysis/data/environmental/HOBO_temperature/20221114_Mahana_3.csv")
mahana.temp.hobo.02 <- read.csv("RAnalysis/data/environmental/HOBO_temperature/20221114_Mahana_2.csv")
```


#bind all data sets together into one data frame

```{r}
temp.data <- rbind(mahana.temp.hobo.02, mahana.temp.hobo.03, mahana.temp.hobo.04, mahana.temp.hobo.05, mahana.temp.hobo.06, mahana.temp.hobo.07, mahana.temp.hobo.08, mahana.temp.hobo.09, mahana.temp.hobo.10, mahana.temp.hobo.11, mahana.temp.hobo.12, mahana.temp.hobo.13)
```


#remove any NA's from the data frame
```{r}
temp.data <- na.omit(temp.data)

view(temp.data)
```

#modify the date and time structure

```{r}
temp.data$date.time <- mdy_hm(temp.data$date.time, quiet=FALSE, tz="UTC", truncated=0) #format date and time
```


#make logger number a factor
```{r}
temp.data$logger.number = as.factor(temp.data$logger.number)
```


#subset temp data for mean across loggers
```{r}
# Separate date and temperature from the "date.temp" column
temp.data.sep <- temp.data %>%
  separate(date.time, into = c("date", "time"), sep = " ")

# Convert Temp to numeric
temp.data.sep$temp <- as.numeric(temp.data.sep$temp)

# Group by Date and calculate the mean temperature for each day
mean_temp_per_day <- temp.data.sep %>%
  group_by(date) %>%
  summarise(mean_temp = mean(temp))

# Assuming your "date" column is not yet of class Date
mean_temp_per_day$date <- as.Date(mean_temp_per_day$date)

filtered_data <- mean_temp_per_day %>%
  filter((date >= as.Date("2022-01-24") & date <= as.Date("2022-10-31")))

filtered_data <- filtered_data %>%
  filter(date != as.Date("2022-08-07"))

```


#plot data by hobo logger per every 4 days no average

```{r}

# plot data for temp change every four days
temp.plot <- filtered_data %>%
  ggplot(aes(x=date, y=mean_temp))+
  geom_line() + 
  geom_point() +
  scale_x_date(date_breaks = "4 day", date_labels = "%m-%d") +
  theme_classic()+
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("Temperature (°C)")+
  xlab("Date")+
  theme(axis.title = element_text(size = 16),   # Adjust the size of x and y axis labels
        axis.text = element_text(size = 16)) ;temp.plot

#ggsave(filename="RAnalysis/output/environmental/20240129_temp_logger_plot.pdf", plot=temp.plot, width=30, height=10, units="in")

ggsave(filename="RAnalysis/output/environmental/20240129_temp_logger_plot.png", plot=temp.plot, width=20, height=10, units="in")
```




# plot temeprature mean sea surface temp per month
```{r}
# Calculate mean sea surface temperature per month
temp.data$month <- factor(month(temp.data$date.time), levels = 1:12, labels = month.abb)  # Extract month from date as factor

#filter to not show november 
temp.dat <- temp.data %>%
  filter(month != "Nov")

#group data by month and summarise for mean temp per month
mean_temp_per_month <- temp.dat %>%
  group_by(month) %>%
  summarise(
    mean_temp = mean(temp, na.rm = TRUE),
    standard_error = sd(temp, na.rm = TRUE) / sqrt(n())) %>%
  arrange(match(month, month.abb))  # Sort the data by month in ascending order

# Plot mean sea surface temperature per month as a dot and line plot
temp.month <- ggplot(mean_temp_per_month, aes(x = as.numeric(month), y = mean_temp)) +
  geom_point(size = 3, color = "black") +
  geom_line(color = "black") +  # Specify the color for the line
  ylab("Mean Sea Surface Temperature (°C)") +
  xlab("Month") +
  theme_classic() +
  scale_x_continuous(breaks = 1:12, labels = month.abb) + # Treat month as continuous
  ylim(NA, 29)  # Adjust the y-axis limits as needed

temp.month

# save mean temp per month plot
#ggsave("RAnalysis/output/environmental/mean.temp.month.pdf", temp.month , width = 10, height = 6)
ggsave("RAnalysis/output/environmental/mean.temp.month.png", temp.month , width = 10, height = 6)


```


# plot temeprature mean sea surface temp per month with variation around
```{r}
# Ensure data only contains January to October
temp.dat <- temp.dat %>%
  filter(month(date.time) %in% 1:10)  # Filter to keep only January to October

# Make smoothed line plot
temp_summary_plot <- temp.dat %>%
  ggplot(aes(x = date.time, y = temp)) + 
  geom_point(colour = "gray", size = 0.5, alpha = 0.05) + 
  geom_smooth(se = FALSE, method = "gam", colour = "black") + 
  ylab("Temperature °C") + 
  xlab("Date") +
  scale_x_datetime(date_breaks = "1 month", date_labels = "%b %Y", 
                   limits = c(as.POSIXct("2022-01-25"), as.POSIXct("2022-10-31"))) +
  theme_classic() + 
  theme(text = element_text(size = 14, colour = "black"), 
    axis.title = element_text(size = 18, colour = "black", face = "bold"), 
    axis.text.x = element_text(size = 14, colour = "black", angle = 45, hjust = 1, vjust = 1),
    legend.title = element_text(size = 18, colour = "black", face = "bold")) + 
  ylim(25, 30);temp_summary_plot

# save mean temp per month plot
ggsave("RAnalysis/output/environmental/temp_summary_plot.png", temp_summary_plot , width = 10, height = 6)

write.csv(temp.data, "gametogenesis_timeseries/gametogenesis_timeseries_analysis/Output/mean_temp_data.csv")


```

##Processing continuous light data for from Odyssey light loggers out at Mahana from March - October

#load light data

```{r}
odyssey.1 <- read.csv("RAnalysis/data/environmental/odyssey_light_loggers/15640_001_003.csv")
odyssey.2 <- read.csv("RAnalysis/data/environmental/odyssey_light_loggers/15641_001_003.csv")

```


```{r}
odyssey.1 <- read.csv("RAnalysis/data/environmental/odyssey_light_loggers/15640_001_003.csv")
odyssey.2 <- read.csv("RAnalysis/data/environmental/odyssey_light_loggers/15641_001_003.csv")

```

#bind all data sets together into one data frame

```{r}
light.data <- rbind(odyssey.1, odyssey.2)
```


#remove any NA's from the data frame
```{r}
light.data <- na.omit(light.data)

view(light.data)
```


```{r}
# Combine Date and Time columns into a new column named DateTime
light.data$DateTime <- with(light.data, {
  dmy_hms(paste(date, time), truncated = 2)  # truncated = 2 allows for two-digit years
})

# Separate date and time from the "date.temp" column
light.data.sep <- light.data %>%
  separate(DateTime, into = c("Date", "Time"), sep = " ")

# Assuming your "Date" column is in a character format
light.data.sep$Date <- as.Date(light.data.sep$Date, format = "%Y-%m-%d")

# remove first two columns
light.data.save <- light.data.sep %>%
  dplyr::select(-date, -time, -raw)

# Group by Date and calculate the mean temperature for each day
mean_light_per_day <- light.data.sep %>%
  group_by(Date) %>%
  summarise(mean_light = mean(calibrated))

# Calculate light per month 
# Convert the Date column to a Date object
light.data.sep$Date <- as.Date(light.data.sep$Date)

# Calculate mean light per month
# Calculate mean light and standard error per month
mean_light_per_month <- light.data.sep %>%
  group_by(month = format(Date, "%Y-%m")) %>%
  summarise(
    mean_light = mean(calibrated, na.rm = TRUE),
    standard_error = sd(calibrated, na.rm = TRUE) / sqrt(n()))

```

```{r}
#compare  light across timeseries
light.plot <- mean_light_per_day %>%
  ggplot(aes(x = Date, y = mean_light)) +
  geom_line() + 
  geom_point() +
  scale_x_date(date_breaks = "4 days", date_labels = "%m-%d") +
  theme_classic() +
  theme(
    axis.text.x = element_text(angle = 90, size = 16),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 16)) +
  xlab("Date") +
  ylab(expression(atop("Photosynthetic Photon Flux Density", (mu*mol~photons~m^{-2}~s^{-1})))); light.plot


#ggsave(filename="RAnalysis/output/environmental/20240129_light_logger_plot.pdf", plot=light.plot, width=30, height=10, units="in")

ggsave(filename="RAnalysis/output/environmental/20240129_light_logger_plot.png", plot=light.plot, width=20, height=10, units="in")

```

# plot light mean light per month
```{r}

# Calculate mean sea surface temperature per month
light.data$month <- factor(month(light.data$DateTime), levels = 1:12, labels = month.abb)  # Extract month from date as factor

#group data by month and summarise for mean temp per month
mean_light_per_month <- light.data %>%
  group_by(month) %>%
  summarise(mean_light = mean(calibrated, na.rm = TRUE)) %>%
  arrange(match(month, month.abb))  # Sort the data by month in ascending order

# Plot mean sea surface temperature per month as a dot and line plot
light.month <- ggplot(mean_light_per_month, aes(x = as.numeric(month), y = mean_light)) +
  geom_point(size = 3, color = "black") +
  geom_line(color = "black") +  # Specify the color for the line
  ylab(expression(atop("Photosynthetic Photon Flux Density", (mu*mol~photons~m^{-2}~s^{-1})))) +
  xlab("Month") +
  theme_classic() +
  scale_x_continuous(breaks = 1:12, labels = month.abb) + # Treat month as continuous
  ylim(NA, 4500)  # Adjust the y-axis limits as needed

light.month

# save mean temp per month plot
#ggsave("RAnalysis/output/environmental/mean.light.month.pdf", light.month , width = 10, height = 6)
ggsave("RAnalysis/output/environmental/mean.light.month.png", light.month , width = 10, height = 6)

```


# plot light per month with variation
```{r}
# Plot mean sea surface temperature per month with variation around the mean
light_summary_plot <- light.data %>%
  ggplot(., aes(x = DateTime, y = calibrated)) + 
  geom_point(colour = "darkgray", size = 0.5, alpha = 0.05) + 
  geom_smooth(aes(), se = FALSE, colour = "black") + 
  ylab(expression(atop("Photosynthetic Photon Flux Density", (mu*mol~photons~m^{-2}~s^{-1})))) +
  xlab("Date") +
  scale_x_datetime(date_breaks = "1 months", date_labels = "%b %Y") +
  theme_classic() + 
  theme(
    text = element_text(size = 12, colour = "black"), 
    axis.title = element_text(size = 14, colour = "black", face = "bold"), 
    axis.text.x = element_text(size = 10, colour = "black", angle = 45, hjust = 1, vjust = 1),
    legend.title = element_text(size = 14, colour = "black", face = "bold")) +
    scale_y_continuous(expand = c(0, 0)) +
    coord_cartesian(ylim = c(1000, 6000))

light_summary_plot

# save mean temp per month plot
#ggsave("RAnalysis/output/environmental/light_summary_plot.pdf", light_summary_plot , width = 10, height = 6)
ggsave("RAnalysis/output/environmental/light_summary_plot.png", light_summary_plot , width = 10, height = 6)

write.csv(light.data, "gametogenesis_timeseries/gametogenesis_timeseries_analysis/Output/mean_light_data.csv")

```

# Make a commbined plot of light and temperature

```{r}
# Convert month from YYYY-MM to just month names (e.g., Jan) in mean_light_per_month
#mean_light_per_month$month <- format(as.Date(paste0(mean_light_per_month$month, "-01")), "%b")

# Join the data frames by month
mean_monthly_data <- full_join(mean_light_per_month, mean_temp_per_month, by = "month")

# Ensure month is a factor for proper ordering in the plot
mean_monthly_data$month <- factor(mean_monthly_data$month,
                                   levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                                              "Jul", "Aug", "Sep", "Oct"))

# Calculate transformed light intensity and temperature for plotting
mean_monthly_data <- mean_monthly_data %>%
    mutate(transformed_light = (mean_light - 2000) / 833 + 26,
        light_se = standard_error / 833,
        transformed_temp = mean_temp,
        temp_se = standard_error)

# Adjusted plotting code with focus on visibility
light.temp <- ggplot(mean_monthly_data, aes(x = month)) +
  
    # Temperature points and error bars
    geom_point(aes(y = transformed_temp, color = "Temperature"), size = 3) +  
    geom_errorbar(aes(y = transformed_temp,
                      ymin = transformed_temp - temp_se,
                      ymax = transformed_temp + temp_se,
                      color = "Temperature"), 
                   width = 0.2) +  
  
    # Light intensity points and error bars
    geom_point(aes(y = transformed_light, color = "Light Intensity"), size = 3) +  
    geom_errorbar(aes(y = transformed_light,
                      ymin = transformed_light - light_se,
                      ymax = transformed_light + light_se,
                      color = "Light Intensity"), 
                   width = 0.2) +  

    # Set up y-axes with adjusted limits
    scale_y_continuous(
        sec.axis = sec_axis(~ (. - 26) * 833 + 2000,
                            name = expression(atop("Photosynthetic Photon Flux Density",
                                                    (mu*mol~photons~m^{-2}~s^{-1})))), 
        name = "Temperature (°C)",
        limits = c(25.5, 29)  # Adjust limits for better visibility
    ) +  

    # Color settings and labels
    scale_color_manual(values = c("Temperature" = "black", "Light Intensity" = "grey")) +
    labs(x = "Month", y = NULL) +
    theme_classic() +
    labs(color = "Environmental Data") +
    theme(legend.position = "none",
          axis.title.y = element_text(color = "black", size = 18),
          axis.text.y = element_text(color = "black", size = 18),
          axis.title.x = element_text(color = "black", size = 18),
          axis.text.x = element_text(color = "black", size = 18),
          axis.title.y.right = element_text(color = "darkgrey", size = 18),
          axis.text.y.right = element_text(color = "darkgrey", size = 18))

light.temp

# save mean temp per month plot
#ggsave("RAnalysis/output/environmental/mean.light.temp.pdf", light.temp , width = 10, height = 6)
ggsave("RAnalysis/output/environmental/mean.light.temp.png", light.temp , width = 10, height = 8)



```

# Making wind speed and precip plost from MCR and adding light and temp separate plots from our data and using PAR to calculate photoperiod

```{r}
# Load MCR data from https://portal.edirepository.org/nis/mapbrowse?packageid=knb-lter-mcr.9.54

mcr <- read.csv("../../../Desktop/Gump_MetData_5_min_average_20251007.csv", stringsAsFactors = FALSE)

mcr <- mcr %>%
  mutate(
    time_local = ymd_hms(time_local),
    time_utc = ymd_hms(time_utc),
    date = as.Date(time_local),
    month = month(date, label = TRUE, abbr = TRUE),
    air_temp_C = na_if(air_temp_C, -999),
    cumulative_rainfall_mm = na_if(cumulative_rainfall_mm, -999),
    wind_U_mps = ifelse(wind_U_mps %in% c(-999, 999), NA, wind_U_mps),
    wind_V_mps = ifelse(wind_V_mps %in% c(-999, 999), NA, wind_V_mps),
    wind_speed = sqrt(wind_U_mps^2 + wind_V_mps^2)
  )

mcr_monthly <- mcr %>%
  group_by(month) %>%
  summarise(
    mean_precip = mean(cumulative_rainfall_mm, na.rm = TRUE),
    se_precip = sd(cumulative_rainfall_mm, na.rm = TRUE) / sqrt(sum(!is.na(cumulative_rainfall_mm))),
    mean_wind = mean(wind_speed, na.rm = TRUE),
    se_wind = sd(wind_speed, na.rm = TRUE) / sqrt(sum(!is.na(wind_speed))),
    .groups = "drop"
  ) %>%
  mutate(month = factor(month, levels = month.abb))

```


```{r}
#PHOTOPERIOD (FROM PAR — KEY STEP): daily light window: We define photoperiod as: first PAR > threshold → last PAR > threshold (use threshold to avoid sensor noise)

threshold <- 5

light.data <- light.data %>%
  mutate(
    DateTime = parse_date_time(DateTime, orders = c("ymd HMS", "ymd HM", "ymd H"), tz = "UTC"),
    date = as.Date(DateTime)
  )

photoperiod_daily <- light.data %>%
  group_by(date) %>%
  summarise(
    first_light = ifelse(
      any(calibrated > threshold, na.rm = TRUE),
      min(DateTime[calibrated > threshold], na.rm = TRUE),
      NA
    ),
    last_light = ifelse(
      any(calibrated > threshold, na.rm = TRUE),
      max(DateTime[calibrated > threshold], na.rm = TRUE),
      NA
    ),
    photoperiod_h = as.numeric(difftime(last_light, first_light, units = "hours")),
    .groups = "drop"
  ) %>%
  filter(!is.na(photoperiod_h) & is.finite(photoperiod_h))

photoperiod_monthly <- photoperiod_daily %>%
  mutate(
    month = factor(month(date, label = TRUE, abbr = TRUE), levels = month.abb)
  ) %>%
  group_by(month) %>%
  summarise(
    mean_photoperiod = mean(photoperiod_h, na.rm = TRUE),
    se_photoperiod = sd(photoperiod_h, na.rm = TRUE) / sqrt(sum(!is.na(photoperiod_h))),
    .groups = "drop"
  ) %>%
  mutate(month = factor(month, levels = month.abb))


```

# Plot temperature, light intensity (PAR), precipitation, wind speed, photoperiod (Van Woesik method)

```{r}

# Make sure month structure is correct for mean monthly data
# Temperature
mean_monthly_data <- mean_monthly_data %>%
  mutate(month = factor(month, levels = month.abb))

month_levels <- month.abb
season_x <- 4.5  # between Apr and May

temp_monthly <- mean_monthly_data %>%
  mutate(month = factor(month, levels = month_levels))

mcr_monthly <- mcr_monthly %>%
  mutate(month = factor(month, levels = month_levels))

photoperiod_monthly <- photoperiod_monthly %>%
  mutate(month = factor(month, levels = month_levels))

p_temp <- ggplot(temp_monthly, aes(x = month, y = mean_temp, group = 1)) +
  geom_line(color = "#4C89C6", linewidth = 1) +
  geom_point(color = "#4C89C6", size = 3) +
  geom_errorbar(aes(ymin = mean_temp - temp_se,
                    ymax = mean_temp + temp_se),
                width = 0.2, color = "#D95F5F") +
  geom_vline(xintercept = season_x, color = "#CC4C4C", linewidth = 1) +
  annotate("text", x = 2, y = -Inf, label = "warm", vjust = -1.2,
           color = "#CC4C4C", fontface = "bold", size = 5) +
  annotate("text", x = 8, y = -Inf, label = "cool", vjust = -1.2,
           color = "#CC4C4C", fontface = "bold", size = 5) +
  theme_classic() +
  labs(title = "Temperature", y = "Temperature (°C)", x = "Month") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 11)
  )

p_temp

p_par <- ggplot(temp_monthly, aes(x = month, y = mean_light, group = 1)) +
  geom_line(color = "#4C89C6", linewidth = 1) +
  geom_point(color = "#4C89C6", size = 3) +
  geom_errorbar(aes(ymin = mean_light - light_se,
                    ymax = mean_light + light_se),
                width = 0.2, color = "#D95F5F") +
  geom_vline(xintercept = season_x, color = "#CC4C4C", linewidth = 1) +
  annotate("text", x = 2, y = -Inf, label = "warm", vjust = -1.2,
           color = "#CC4C4C", fontface = "bold", size = 5) +
  annotate("text", x = 8, y = -Inf, label = "cool", vjust = -1.2,
           color = "#CC4C4C", fontface = "bold", size = 5) +
  theme_classic() +
  labs(title = "PAR", y = expression(atop("Photosynthetic Photon Flux Density",
                                           (mu*mol~photons~m^{-2}~s^{-1}))),
       x = "Month") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 11)
  )

p_par

p_precip <- ggplot(mcr_monthly, aes(x = month, y = mean_precip, group = 1)) +
  geom_line(color = "#4C89C6", linewidth = 1) +
  geom_point(color = "#4C89C6", size = 3) +
  geom_errorbar(aes(ymin = mean_precip - se_precip,
                    ymax = mean_precip + se_precip),
                width = 0.2, color = "#D95F5F") +
  geom_vline(xintercept = season_x, color = "#CC4C4C", linewidth = 1) +
  annotate("text", x = 2, y = -Inf, label = "warm", vjust = -1.2,
           color = "#CC4C4C", fontface = "bold", size = 5) +
  annotate("text", x = 8, y = -Inf, label = "cool", vjust = -1.2,
           color = "#CC4C4C", fontface = "bold", size = 5) +
  theme_classic() +
  labs(title = "Precipitation", y = "mm/mo", x = "Month") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 11)
  )

p_precip

p_wind <- ggplot(mcr_monthly, aes(x = month, y = mean_wind, group = 1)) +
  geom_line(color = "#4C89C6", linewidth = 1) +
  geom_point(color = "#4C89C6", size = 3) +
  geom_errorbar(aes(ymin = mean_wind - se_wind,
                    ymax = mean_wind + se_wind),
                width = 0.2, color = "#D95F5F") +
  geom_vline(xintercept = season_x, color = "#CC4C4C", linewidth = 1) +
  annotate("text", x = 2, y = -Inf, label = "warm", vjust = -1.2,
           color = "#CC4C4C", fontface = "bold", size = 5) +
  annotate("text", x = 8, y = -Inf, label = "cool", vjust = -1.2,
           color = "#CC4C4C", fontface = "bold", size = 5) +
  theme_classic() +
  labs(title = "Wind Speed", y = "m/s", x = "Month") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.title = element_text(size = 14),
    axis.text = element_text(size = 11)
  )

p_wind 

season_x <- 2.5  # between Apr and May

p_photo <- ggplot(photoperiod_monthly, aes(x = month, y = mean_photoperiod, group = 1)) +
  geom_line(color = "#4C89C6", linewidth = 1) +
  geom_point(color = "#4C89C6", size = 3) +
  geom_errorbar(aes(ymin = mean_photoperiod - se_photoperiod,
                    ymax = mean_photoperiod + se_photoperiod),
                width = 0.2, color = "#D95F5F") +
  geom_vline(xintercept = season_x, color = "#CC4C4C", linewidth = 1) +
  annotate("text", x = 1.5, y = -Inf, label = "warm", vjust = -1.2,
           color = "#CC4C4C", fontface = "bold", size = 5) +
  annotate("text", x = 5.5, y = -Inf, label = "cool", vjust = -1.2,
           color = "#CC4C4C", fontface = "bold", size = 5) +
  theme_classic() +
  labs(title = "Photoperiod", y = "hr", x = "Month")

p_photo
  
# FINAL FIGURE 
bottom_row <- p_photo + plot_spacer()

(p_temp | p_par) /
(p_precip | p_wind) /
bottom_row
```

# Bring in histological data to make figures next
```{r}
oocyte_abun_data <- read.csv("gametogenesis_timeseries/gametogenesis_timeseries_analysis/Output/oocyte_abun_data.csv")
sperm_abun_data  <- read.csv("gametogenesis_timeseries/gametogenesis_timeseries_analysis/Output/sperm_abun_data.csv")
oogen_size_data  <- read.csv("gametogenesis_timeseries/gametogenesis_timeseries_analysis/Output/oogen_size_data.csv")
sperm_size_data  <- read.csv("gametogenesis_timeseries/gametogenesis_timeseries_analysis/Output/sperm_size_data.csv")

# combine abundance data for gametes
gamete_monthly <- oocyte_abun_data %>%
  group_by(timepoint) %>%
  summarise(eggs_pct = sum(freq[Stage != "Stage 0"], na.rm = TRUE), .groups = "drop") %>%
  left_join(
    sperm_abun_data %>%
      group_by(timepoint) %>%
      summarise(sperm_pct = sum(freq[Stage != "Stage 0"], na.rm = TRUE), .groups = "drop"),
    by = "timepoint"
  )
```


# Make final figure temperature, windspeed, and light for insolation and photoperiod and then the z score of each of these variables on y and x has months from december 2021 to november 2022, then other y axes is % of individuals with gametes the bar columns would then have eggs and sperm percentages (not stages)

```{r}
# monthly labels for final figure
month_levels <- c(
  "Dec 2021","Jan 2022","Feb 2022","Mar 2022","Apr 2022","May 2022",
  "Jun 2022","Jul 2022","Aug 2022","Sep 2022","Oct 2022","Nov 2022"
)

month_map <- c(
  "Dec" = "Dec 2021",
  "Jan" = "Jan 2022",
  "Feb" = "Feb 2022",
  "Mar" = "Mar 2022",
  "Apr" = "Apr 2022",
  "May" = "May 2022",
  "Jun" = "Jun 2022",
  "Jul" = "Jul 2022",
  "Aug" = "Aug 2022",
  "Sep" = "Sep 2022",
  "Oct" = "Oct 2022",
  "Nov" = "Nov 2022"
)

# histology abundance summaries
oocyte_abun_data <- read.csv("gametogenesis_timeseries/gametogenesis_timeseries_analysis/Output/oocyte_abun_data.csv")
sperm_abun_data  <- read.csv("gametogenesis_timeseries/gametogenesis_timeseries_analysis/Output/sperm_abun_data.csv")

gamete_monthly <- gamete_monthly %>%
  mutate(month = case_when(
    timepoint == "December"  ~ "Dec 2021",
    timepoint == "January"   ~ "Jan 2022",
    timepoint == "February"  ~ "Feb 2022",
    timepoint == "March"     ~ "Mar 2022",
    timepoint == "April"     ~ "Apr 2022",
    timepoint == "May"       ~ "May 2022",
    timepoint == "June"      ~ "Jun 2022",
    timepoint == "July"      ~ "Jul 2022",
    timepoint == "August"    ~ "Aug 2022",
    timepoint == "September" ~ "Sep 2022",
    timepoint == "October"   ~ "Oct 2022",
    timepoint == "November"  ~ "Nov 2022",
    TRUE ~ NA_character_
  )) %>%
  select(month, eggs_pct, sperm_pct)

# environmental summaries must also use the same month labels
temp_monthly <- mean_temp_per_month %>%
  mutate(
    month = recode(as.character(month), !!!month_map),
    month = factor(month, levels = month_levels),
    temp_z = as.numeric(scale(mean_temp))
  )

light_monthly <- mean_light_per_month %>%
  mutate(
    month = recode(as.character(month), !!!month_map),
    month = factor(month, levels = month_levels),
    light_z = as.numeric(scale(mean_light))
  )

photoperiod_monthly <- photoperiod_monthly %>%
  mutate(
    month = as.character(month),
    month = recode(month, !!!month_map),
    month = factor(month, levels = month_levels),
    photo_z = as.numeric(scale(mean_photoperiod))
  )

mcr_monthly2 <- mcr %>%
  group_by(month) %>%
  summarise(
    mean_precip = mean(cumulative_rainfall_mm, na.rm = TRUE),
    se_precip = sd(cumulative_rainfall_mm, na.rm = TRUE) / sqrt(sum(!is.na(cumulative_rainfall_mm))),
    mean_wind = mean(wind_speed, na.rm = TRUE),
    se_wind = sd(wind_speed, na.rm = TRUE) / sqrt(sum(!is.na(wind_speed))),
    .groups = "drop"
  ) %>%
  mutate(
    month = factor(month, levels = month.abb),
    wind_z = as.numeric(scale(mean_wind))
  )

gamete_monthly <- gamete_monthly %>%
  mutate(month = as.character(month))

temp_monthly <- temp_monthly %>%
  mutate(month = as.character(month))

mcr_monthly2 <- mcr_monthly2 %>%
  mutate(month = as.character(month))

mcr_monthly2 <- mcr_monthly2 %>%
  mutate(
    month = case_when(
      month == "Jan" ~ "Jan 2022",
      month == "Feb" ~ "Feb 2022",
      month == "Mar" ~ "Mar 2022",
      month == "Apr" ~ "Apr 2022",
      month == "May" ~ "May 2022",
      month == "Jun" ~ "Jun 2022",
      month == "Jul" ~ "Jul 2022",
      month == "Aug" ~ "Aug 2022",
      month == "Sep" ~ "Sep 2022",
      month == "Oct" ~ "Oct 2022",
      month == "Nov" ~ "Nov 2022",
      month == "Dec" ~ "Dec 2021"
    )
  )

light_monthly <- light_monthly %>%
  mutate(month = as.character(month))

photoperiod_monthly <- photoperiod_monthly %>%
  mutate(month = as.character(month))

env_z <- gamete_monthly %>%
  left_join(temp_monthly %>% select(month, temp_z), by = "month") %>%
  left_join(mcr_monthly2 %>% select(month, wind_z), by = "month") %>%
  left_join(light_monthly %>% select(month, light_z), by = "month") %>%
  left_join(photoperiod_monthly %>% select(month, photo_z), by = "month")

p_bottom <- ggplot(env_z, aes(x = month)) +
  geom_col(aes(y = eggs_pct, fill = "Eggs"), width = 0.7) +
  geom_col(aes(y = sperm_pct, fill = "Sperm"), width = 0.7, position = "stack") +
  geom_line(aes(y = temp_z, color = "Temperature", group = 1), linewidth = 1) +
  geom_line(aes(y = wind_z, color = "Wind speed", group = 1), linewidth = 1) +
  geom_line(aes(y = light_z, color = "Light", group = 1), linewidth = 1) +
  geom_line(aes(y = photo_z, color = "Photoperiod", group = 1), linewidth = 1) +
  scale_y_continuous(
    name = "Z-score",
    sec.axis = sec_axis(~ ., name = "% individuals with gametes")
  ) +
  scale_fill_manual(values = c("Eggs" = "#4C89C6", "Sperm" = "#808080")) +
  scale_color_manual(values = c(
    "Temperature" = "#1b9e77",
    "Wind speed" = "#d95f02",
    "Light" = "#7570b3",
    "Photoperiod" = "#e7298a"
  )) +
  theme_classic() +
  labs(x = "Month", fill = NULL, color = NULL) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top")

(p_temp | p_wind) / (p_par | p_photo) / p_bottom
```

