---
title: "Gamete Abundance Analysis"
author: "daniellembecker"
date: "2024-03-31"
output: html_document
---

This script will analyze gamete abundance in Acropora pulchra over time. The data is separated into timepointly timepoints and spermatogenesis and oogenesis.

## Load packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("tidyverse")
library("gridExtra")
library("ghibli")
library("PNWColors")
library("lubridate")
```

## Load data & select specific columns 
```{r}
#load data files
abundance <- read.csv("histology/data/histological_data.csv", header = T, na.strings = c("", "NA"))

# organize data make timepoints factors with levels in order for plots and calculate frequency
df_relative_abundance <- abundance %>%
  count(colony_id, replicate_id, site, timepoint, Stage, gonad.type, name = "stage_count") %>%
  group_by(colony_id, replicate_id, site, timepoint, gonad.type) %>%
  mutate(grand_tots = sum(stage_count)) %>%
  mutate(freq = stage_count / grand_tots)

# change the freq to 0 whenever the corresponding Stage column has NA
df_relative_abundance <- df_relative_abundance %>%
  mutate(freq = ifelse(is.na(Stage), 0, freq),
    grand_tots = ifelse(is.na(Stage), 0, grand_tots),
    stage_count = ifelse(is.na(Stage), 0, stage_count))

# creates a new column in the df_relative_abundance dataframe called group, which is a combination of the timepoint and colony_id columns
df_relative_abundance$group <- paste0(df_relative_abundance$timepoint,"_", df_relative_abundance$colony_id)

# convert the Stage column in the df_relative_abundance dataframe to a factor with specified levels
df_relative_abundance$Stage <- factor(df_relative_abundance$Stage, levels = c("StageI","StageII","StageIII","StageIV", "StageV"))

```

# Plot Oogenesis abundance plots
```{r}
# select just oocytes, remove any spermatocyte data
oogen_df_relative_abundance <- df_relative_abundance %>%
  filter(gonad.type != "spermatocyte")

# Reorder the levels of the 'timepoint' variable from December to October
oogen_df_relative_abundance$timepoint <- factor(oogen_df_relative_abundance$timepoint, 
                                                 levels = c("December", "January", "February", "March", "April", "May", "June", "July", "August", "September", "October"),
                                                 ordered = TRUE)

# Plot without geom_text for each stacked plot
oogen <- oogen_df_relative_abundance %>% 
  ggplot(aes(x = replicate_id, y = freq, fill = Stage)) +
  geom_col() +
  facet_wrap(~ timepoint, ncol = 11) +
  theme_bw() +
  ylab("Proportion of Oocytes") +
  xlab("Colony ID") +
  theme(axis.text.x = element_text(angle = 90, size = 18),  # Increase size of x-axis text
        axis.text.y = element_text(size = 18),  # Increase size of y-axis text
        legend.position = "top",
        legend.title = element_blank(),
        text = element_text(size = 18)) +  # Increase size of all other text elements
  scale_y_continuous(expand = c(0, 0)) +
  guides(color = guide_legend(ncol = length(unique(oogen_df_relative_abundance$Stage)), byrow = TRUE), fill = guide_legend(ncol = length(unique(oogen_df_relative_abundance$Stage)), byrow = TRUE))

oogen

# Add grand_tots value on the x-axis
oogen2 <- oogen +
  geom_text(aes(x = replicate_id, y = 0, label = grand_tots), vjust = -0.5, size = 5, color = "black")

oogen2

# save oogenesis plot
ggsave("histology/output/oogenesis_abundance.pdf", oogen2 , width = 13, height = 8)
ggsave("histology/output/oogenesis_abundance.jpg", oogen2 , width = 13, height = 8)


```

# Plot Spermatogenesis abundance plots
```{r}
spermato_df_relative_abundance <- df_relative_abundance %>%
  filter(!gonad.type=="oocyte")

spermato <- spermato_df_relative_abundance %>% ggplot(aes(x=colony_id, y=freq, group=stage, color=stage, fill=stage)) +
  geom_col()+
  #geom_text(aes(label=grand.tots),color="black", size=1.5, vjust = 0.1)+
  facet_wrap(c("site", "timepoint"), ncol=11)+
  theme_bw()+
  ylab("Proportion of Sperm")+
  theme(axis.text.x = element_text(angle = 90))


```

# Arrange and save abundance plots
```{r}

#plots <- grid.arrange(oogen, spermato, nrow = 2)


```

# Plot Oogenesis size by replicate 
```{r}

# select for grand mean size dataframe
oogen_size <- abundance %>%
  dplyr::select(replicate_id, colony_id, timepoint, gonad.type, geometric.mean, Stage) 

# select just oocytes, remove any spermatocyte data
oogen_size <- oogen_size %>%
  filter(gonad.type != "spermatocyte")

# count number of oocytes per timmepoint for count on x axis
oogen_size <- oogen_size %>%
  group_by(colony_id, timepoint) %>%
  mutate(count = n()) %>%
  ungroup()

# Reorder the levels of the 'timepoint' variable from December to October
oogen_size$timepoint <- factor(oogen_size$timepoint, 
                                                 levels = c("December", "January", "February", "March", "April", "May", "June", "July", "August", "September", "October"),
                                                 ordered = TRUE)

# Calculate mean oocyte size per month per replicate
mean_size <- oogen_size %>%
  group_by(timepoint, replicate_id, count) %>%
  summarise(mean_size = mean(geometric.mean, na.rm = TRUE))

# Plot with reordered months and size and by replicate
oogen.size.plot <- mean_size %>% 
  ggplot(aes(x = replicate_id, y = mean_size)) +
  geom_col() +
  facet_wrap(~ timepoint, ncol = 11) +
  theme_bw() +
  ylab("Oocyte Diameter (microns)") +
  xlab("Colony ID") +
  theme(axis.text.x = element_text(angle = 90, size = 18),  # Increase size of x-axis text
        axis.text.y = element_text(size = 18),  # Increase size of y-axis text
        legend.position = "top",
        legend.title = element_blank(),
        text = element_text(size = 18)) +  # Increase size of all other text elements
  ylab(expression("Mean oocyte diameter ("*mu*"m)")) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 400)) + # Ensure the y-axis starts at 0
  guides(fill = guide_legend(ncol = length(unique(oogen_size$Stage)), byrow = TRUE))  # Set ncol and byrow for the legend # Remove legend titles

oogen.size.plot

# Add count value on the x-axis
oogen.size.plot2 <- oogen.size.plot +
  geom_text(aes(x = replicate_id, y = 0, label = count), vjust = -0.5, size = 5, color = "black")

oogen.size.plot2

# save oogenesis size plots
ggsave("histology/output/oogenesis_size.pdf", oogen.size.plot2 , width = 13, height = 8)
ggsave("histology/output/oogenesis_size.jpg", oogen.size.plot2 , width = 13, height = 8)


```

# Plot oogenesis mean size by month

```{r}
# Calculate mean oocyte size per month per replicate
mean_size_month <- oogen_size %>%
  group_by(timepoint) %>%
  summarise(mean_size = mean(geometric.mean, na.rm = TRUE),
            std_error = sd(geometric.mean, na.rm = TRUE) / sqrt(n())) 

# Abbreviate month names
mean_size_month$abbreviated_timepoint <- month.abb[match(mean_size_month$timepoint, month.name)]

# Reorder the levels of the 'timepoint' variable from December to October
mean_size_month$abbreviated_timepoint <- factor(mean_size_month$abbreviated_timepoint, 
                                        levels = c("Dec", "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct"), ordered = TRUE)

#use PNW color palette
pal <- pnw_palette("Moth",12)

# Plot with Studio Ghibli color palette
oogen.size.month <- mean_size_month %>% 
  ggplot(aes(x = abbreviated_timepoint, y = mean_size)) +
  geom_col() +
  geom_errorbar(aes(ymin = mean_size - std_error, ymax = mean_size + std_error), width = 0.2, color = "black") +  # Add error bars
  #scale_fill_manual(values=pal)  + #specify PNW colors
  theme_classic() +
  ylab("Oocyte Diameter (microns)") +
  xlab("Timepoint") +
  theme(axis.text.x = element_text(angle = 0, size = 18),  # Increase size of x-axis text
        axis.text.y = element_text(size = 18),  # Increase size of y-axis text
        legend.position = "none",  # Remove legend
        text = element_text(size = 18)) +  # Increase size of all other text elements
  ylab(expression("Mean oocyte diameter ("*mu*"m)")) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 410))

oogen.size.month

# save oogenesis size per month plots
ggsave("histology/output/oogen.size.month.pdf", oogen.size.month , width = 13, height = 8)
ggsave("histology/output/oogen.size.month.jpg", oogen.size.month , width = 13, height = 8)

# select only month and temperature for temp data (environmental script must be run first)
temperature <- mean_monthly_data %>% 
  dplyr::select(month, mean_temp) 

# Rename month to timepoint to overlay with plot
temperature <- temperature %>%
  rename(timepoint = month)

# Calculate temp_adjustment
temp_adjustment <- (29 - 26) / (max(temperature$mean_temp) - min(temperature$mean_temp))

# Plot with temperature overlay and dual y-axis
oogen.size.temp <- oogen.size.month + 
  geom_line(data = temperature, aes(x = timepoint, y = (mean_temp - 26) * 384 / 3 + 26, color = "Temperature", group = 1)) +
  geom_point(data = temperature, aes(x = timepoint, y = (mean_temp - 26) * 384 / 3 + 26, color = "Temperature")) +
  scale_color_manual(values = c("Temperature" = "black")) +  # Adjust color if needed
  theme(axis.text.x = element_text(angle = 0, size = 18),  # Increase size of x-axis text
        axis.text.y = element_text(size = 18),  # Increase size of y-axis text
        legend.position = "none",  # Remove legend
        text = element_text(size = 18)) +  # Increase size of all other text elements
  scale_y_continuous(sec.axis = sec_axis(~ (. - 26) / 384 * 3 + 26, name = "Temperature (°C)"), expand = c(0, 0), limits = c(0, 410))

oogen.size.temp

# save oogenesis size plots with temperature overlayed
ggsave("histology/output/oogen.size.temp.pdf", oogen.size.temp , width = 13, height = 8)
ggsave("histology/output/oogen.size.temp.jpg", oogen.size.temp , width = 13, height = 8)



```






# Plot Spermatogenesis size plots
```{r}

# select for grand mean size dataframe
spermato_size <- abundance %>%
  dplyr::select(replicate_id, colony_id, timepoint, gonad.type, geometric.mean, Stage) 

spermato_size <- spermato_size %>%
  group_by(colony_id, timepoint) %>%
  mutate(count = n()) %>%
  ungroup()

# Reorder the levels of the 'timepoint' variable from December to October
spermato_size$timepoint <- factor(spermato_size$timepoint, 
                                                 levels = c("December", "January", "February", "March", "April", "May", "June", "July", "August", "September", "October"),
                                                 ordered = TRUE)

# Calculate mean oocyte size per month
mean_size <- oogen_size %>%
  group_by(timepoint, replicate_id, count) %>%
  summarise(mean_size = mean(geometric.mean, na.rm = TRUE))

# Plot with reordered months and size
oogen.size.plot <- mean_size %>% 
  ggplot(aes(x = replicate_id, y = mean_size)) +
  geom_col() +
  facet_wrap(~ timepoint, ncol = 11) +
  theme_bw() +
  ylab("Oocyte Diameter (microns)") +
  xlab("Colony ID") +
  theme(axis.text.x = element_text(angle = 90, size = 18),  # Increase size of x-axis text
        axis.text.y = element_text(size = 18),  # Increase size of y-axis text
        legend.position = "top",
        legend.title = element_blank(),
        text = element_text(size = 18)) +  # Increase size of all other text elements
  ylab(expression("Mean oocyte diameter ("*mu*"m)")) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 400)) + # Ensure the y-axis starts at 0
  guides(fill = guide_legend(ncol = length(unique(oogen_size$Stage)), byrow = TRUE))  # Set ncol and byrow for the legend # Remove legend titles

oogen.size.plot

# Add count value on the x-axis
oogen.size.plot2 <- oogen.size.plot +
  geom_text(aes(x = replicate_id, y = 0, label = count), vjust = -0.5, size = 5, color = "black")

oogen.size.plot2

# save oogenesis size plots
ggsave("histology/output/oogenesis_size.pdf", oogen.size.plot2 , width = 13, height = 8)
ggsave("histology/output/oogenesis_size.jpg", oogen.size.plot2 , width = 13, height = 8)


```


# Arrange and save size plots
```{r}
# save oogenesis plot

#plots <- grid.arrange(oogen, spermato, nrow = 2)


```