---
title: "Gamete Abundance Analysis"
author: "daniellembecker"
date: "2024-03-31"
output: html_document
editor_options: 
  chunk_output_type: console
---

This script will analyze gamete abundance in Acropora pulchra over time. The data is separated into timepoints and spermatogenesis and oogenesis.

## Load packages 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("tidyverse")
library("betareg")
library("glmmTMB")
library("gridExtra")
library("ghibli")
library("lme4")
library("nnet")
library("car")
library("PNWColors")
library("lubridate")
library("lsmeans")
library("multcomp")
library("patchwork")

```

## Load data & arrange for relative abundance 
```{r}
#load data files
abundance <- read.csv("histology/data/histological_data.csv", header = T, na.strings = c("", "NA"))

# organize data make timepoints factors with levels in order for plots and calculate frequency
df_relative_abundance <- abundance %>%
  count(colony_id, replicate_id, site, timepoint, Stage, gonad.type, patch, name = "stage_count") %>%
  group_by(colony_id, replicate_id, site, timepoint, gonad.type) %>%
  mutate(grand_tots = sum(stage_count)) %>%
  mutate(freq = stage_count / grand_tots)

# change the freq to 0 whenever the corresponding Stage column has NA
df_relative_abundance <- df_relative_abundance %>%
  mutate(freq = ifelse(is.na(Stage), 0, freq),
    grand_tots = ifelse(is.na(Stage), 0, grand_tots),
    stage_count = ifelse(is.na(Stage), 0, stage_count))

# creates a new column in the df_relative_abundance dataframe called group, which is a combination of the timepoint and colony_id columns
df_relative_abundance$group <- paste0(df_relative_abundance$timepoint,"_", df_relative_abundance$colony_id)

# convert the Stage column in the df_relative_abundance dataframe to a factor with specified levels
df_relative_abundance$Stage <- factor(df_relative_abundance$Stage, levels = c("Stage I","Stage II","Stage III","Stage IV", "Stage V"))


```

# Plot Oogenesis abundance plot by colony
```{r}
# select just oocytes, remove any spermatocyte data
oogen_df_relative_abundance <- df_relative_abundance %>%
  filter(gonad.type != "spermatocyte")

# Reorder the levels of the 'timepoint' variable from December to October
oogen_df_relative_abundance$timepoint <- factor(oogen_df_relative_abundance$timepoint, 
                                                 levels = c("December", "January", "February", "March", "April", "May", "June", "July", "August", "September", "October"),
                                                 ordered = TRUE)

# Plot without geom_text for each stacked plot
oogen <- oogen_df_relative_abundance %>% 
  ggplot(aes(x = replicate_id, y = freq, fill = Stage)) +
  geom_col() +
  facet_wrap(~ timepoint, ncol = 11) +
  theme_bw() +
  ylab("Relative Abundance of Oocytes") +
  xlab("Colony ID") +
  theme(axis.text.x = element_text(angle = 90, size = 18),  # Increase size of x-axis text
        axis.text.y = element_text(size = 18),  # Increase size of y-axis text
        legend.position = "top",
        legend.title = element_blank(),
        text = element_text(size = 18)) +  # Increase size of all other text elements
  scale_y_continuous(expand = c(0, 0)) +
  guides(color = guide_legend(ncol = length(unique(oogen_df_relative_abundance$Stage)), byrow = TRUE), fill = guide_legend(ncol = length(unique(oogen_df_relative_abundance$Stage)), byrow = TRUE))

oogen

# Add grand_tots value on the x-axis
oogen2 <- oogen +
  geom_text(aes(x = replicate_id, y = 0, label = grand_tots), vjust = -0.5, size = 5, color = "black")

oogen2

# save oogenesis plot
ggsave("histology/output/oogenesis_abundance_percolony.png", oogen2 , width = 13, height = 8)


```


# Plot Spermatogenesis abundance plot by colony
```{r}
spermato_df_relative_abundance <- df_relative_abundance %>%
  filter(!gonad.type=="oocyte")

# Reorder the levels of the 'timepoint' variable from December to October
spermato_df_relative_abundance$timepoint <- factor(spermato_df_relative_abundance$timepoint, 
                                                 levels = c("December", "January", "February", "March", "April", "May", "June", "July", "August", "September", "October"),
                                                 ordered = TRUE)

# Plot without geom_text for each stacked plot
spermato <- spermato_df_relative_abundance %>% 
  ggplot(aes(x = replicate_id, y = freq, fill = Stage)) +
  geom_col() +
  facet_wrap(~ timepoint, ncol = 11) +
  theme_bw() +
  ylab("Relative Abundance of Spermaries") +
  xlab("Colony ID") +
  theme(axis.text.x = element_text(angle = 90, size = 18),  # Increase size of x-axis text
        axis.text.y = element_text(size = 18),  # Increase size of y-axis text
        legend.position = "top",
        legend.title = element_blank(),
        text = element_text(size = 18)) +  # Increase size of all other text elements
  scale_y_continuous(expand = c(0, 0)) +
  guides(color = guide_legend(ncol = length(unique(spermato_df_relative_abundance$Stage)), byrow = TRUE), fill = guide_legend(ncol = length(unique(spermato_df_relative_abundance$Stage)), byrow = TRUE))

spermato

# Add grand_tots value on the x-axis
spermato2 <- spermato +
  geom_text(aes(x = replicate_id, y = 0, label = grand_tots), vjust = -0.5, size = 5, color = "black")

spermato2

# save oogenesis plot
ggsave("histology/output/spermatogenesis_abundance_bycolony.png", spermato2 , width = 13, height = 8)



```


# Plot relative abundance per stage per timepoint for oogenesis 
```{r}
# Replace NA in Stage with "Stage 0"
abundance_clean <- abundance %>%
  mutate(Stage = ifelse(is.na(Stage), "Stage 0", as.character(Stage)))

# Filter for oocytes and calculate total counts
rel.abun <- abundance_clean %>%
  filter(gonad.type == "oocyte") %>%
  count(site, timepoint, Stage, gonad.type, name = "stage_count") %>%
  group_by(timepoint) %>%
  mutate(grand_tots = sum(stage_count, na.rm = TRUE)) %>%
  ungroup()

# Calculate frequency for each stage
rel.abun <- rel.abun %>%
  mutate(freq = stage_count / grand_tots)

# Calculate the "No Data" frequency for Stage 0
oocyte.abund.all <- rel.abun %>%
  group_by(timepoint) %>%
  mutate(freq_no_data = 1 - sum(freq[Stage != "Stage 0"], na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(freq = ifelse(Stage == "Stage 0", freq_no_data, freq)) %>%
  dplyr::select(timepoint, Stage, freq) %>%
  distinct()

# Ensure that Stage is a factor and includes all levels, even if missing
oocyte.abund.all$Stage <- factor(oocyte.abund.all$Stage, 
                                                   levels = c("Stage 0", "Stage I", "Stage II", "Stage III", "Stage IV", "Stage V"))

# Reorder the levels of the 'timepoint' variable
oocyte.abund.all$timepoint <- factor(oocyte.abund.all$timepoint, 
                                                       levels = c("December", "January", "February", "March", "April", "May", "June", "July", "August", "September", "October"),
                                                       ordered = TRUE)

# Define color gradient
color_gradient <- c("Stage 0" = "#D3D3D3", "Stage I" = "#EAE2F8", "Stage II" = "#D0A6F7", 
                    "Stage III" = "#B76CF7", "Stage IV" = "#8E2BEA", "Stage V" = "#3B1E8F")

# Plot for Oocytes with gradient color scheme
oocytes_plot <- oocyte.abund.all %>%
  ggplot(aes(x = timepoint, y = freq, fill = Stage)) +
  geom_col(position = "stack") +  # Stack bars to show relative abundance
  scale_fill_manual(values = color_gradient) +
  theme_bw() +
  ylab("Relative Abundance of Oocytes") +
  theme(axis.text.x = element_blank(),  # Remove x-axis text from oocytes plot
        axis.text.y = element_text(size = 20),  # Increase size of y-axis text
        axis.title.x = element_blank(),  # Remove x-axis label
        axis.title.y = element_text(size = 22),
        text = element_text(size = 22),
        legend.position = "top",  # Place legend at the top
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +  # Remove grid lines for clarity
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)); oocytes_plot

# Save the oocytes plot
ggsave("histology/output/oocytes_relative_abundance.png", oocytes_plot, width = 13, height = 8)


# save dataframe for downstream figure
write.csv(oocyte.abund.all, "gametogenesis_timeseries/gametogenesis_timeseries_analysis/Output/oocyte_abun_data.csv")


```

# Plot relative abundance per stage per timepoint for spermatogenesis 
```{r}
# Filter for spermatocytes and calculate total counts
rel.abun_spermatocyte <- abundance_clean %>%
  filter(gonad.type == "spermatocyte") %>%
  count(site, timepoint, Stage, gonad.type,name = "stage_count") %>%
  group_by(timepoint) %>%
  mutate(grand_tots = sum(stage_count, na.rm = TRUE)) %>%
  ungroup()

# Calculate frequency for each stage
rel.abun_spermatocyte <- rel.abun_spermatocyte %>%
  mutate(freq = stage_count / grand_tots)

# Ensure all stages including "No Data in Slides" Stage 0 are present
rel_full_spermatocyte <- rel.abun_spermatocyte %>%
  complete(timepoint, Stage = c("Stage I", "Stage II", "Stage III", "Stage IV", "Stage V", "Stage 0"), fill = list(freq = 0)) %>%
  group_by(timepoint) %>%
  mutate(freq = ifelse(is.na(freq), 0, freq)) %>%
  ungroup()

# Calculate the "No Data" Stage 0 frequency
sperm_abun <- rel_full_spermatocyte %>%
  group_by(timepoint) %>%
  mutate(freq_no_data = 1 - sum(freq[Stage != "Stage 0"], na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(freq = ifelse(Stage == "Stage 0", freq_no_data, freq)) %>%
  dplyr::select(timepoint, Stage, freq) %>%
  distinct()

# Reorder the levels of the 'timepoint' variable
sperm_abun$timepoint <- factor(sperm_abun$timepoint, 
                                                   levels = c("December", "January", "February", "March", "April", "May", "June", "July", "August", "September", "October"),
                                                   ordered = TRUE)

# Define color gradient
color_gradient <- c("Stage I" = "#EAE2F8", "Stage II" = "#D0A6F7", "Stage III" = "#B76CF7", "Stage IV" = "#8E2BEA", "Stage V" = "#3B1E8F", "Stage 0" = "#D3D3D3")

# Plot spermatocytes per month abundance
spermatocytes_plot <- sperm_abun %>%
  ggplot(aes(x = timepoint, y = freq, fill = Stage)) +
  geom_col(position = "stack") +  # Stack bars to show relative abundance
  scale_fill_manual(values = color_gradient) +
  theme_bw() +
  ylab("Relative Abundance of Spermaries") +
  xlab("Timepoint") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 20),  # Rotate x-axis labels 45 degrees and increase size
        axis.text.y = element_text(size = 20),  # Increase size of y-axis text
        axis.title.y = element_text(size = 22),
        axis.title.x = element_text(size = 22),
        text = element_text(size = 22),
        legend.position = ("none"),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +  # Remove grid lines for clarity
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)); spermatocytes_plot


# Save the spermatocyte plot
ggsave("histology/output/spermatocyte_relative_abundance.png", spermatocytes_plot, width = 13, height = 8)


# save dataframe for downstream figure
write.csv(sperm_abun, "gametogenesis_timeseries/gametogenesis_timeseries_analysis/Output/sperm_abun_data.csv")


```

# Combine relative abundance plots with patchwork
```{r}
# Combine plots using patchwork
combined_plot <- (oocytes_plot / spermatocytes_plot) +
  plot_annotation(tag_levels = 'A',  # Automatically add tags 'A', 'B', etc.
                  theme = theme(plot.tag = element_text(size = 24, face = "bold"))) &
  theme(legend.position = "top")  # Position the legend at the top
combined_plot <- combined_plot + plot_layout(guides = "collect");combined_plot  # Collect legends into a single legend

# Save the combined plot
ggsave("histology/output/combined_abundance_plots.png", combined_plot, width = 20, height = 15)
```



# Conduct univariate analyses and plot oocyte size per month
```{r}
# select for grand mean size dataframe
oogen_size <- abundance %>%
  dplyr::select(replicate_id, colony_id, timepoint, gonad.type, geometric.mean, Stage, patch)

# select just oocytes, remove any spermatocyte data
oogen_size <- oogen_size %>%
  filter(gonad.type != "spermatocyte")%>%
  filter(geometric.mean != 0)

# count number of oocytes per timmepoint for count on x axis
oogen_size <- oogen_size %>%
  group_by(colony_id, timepoint) %>%
  mutate(count = if (all(is.na(geometric.mean) | geometric.mean == 0)) 0 else n()) %>%
  ungroup()

# Reorder the levels of the 'timepoint' variable from December to October
oogen_size$timepoint <- factor(oogen_size$timepoint, 
                                                 levels = c("December", "January", "February", "March", "April", "May", "June", "July", "August", "September", "October"),
                                                 ordered = TRUE)

# Calculate mean oocyte size per month per replicate (keep this step as needed for other purposes)
mean_size_month <- oogen_size %>%
  group_by(timepoint) %>%
  summarise(mean_size = mean(geometric.mean, na.rm = TRUE),
            std_error = sd(geometric.mean, na.rm = TRUE) / sqrt(n())) 

# compare if colony_id as random effect improves model
# Model with colony as a random effect
model_with_colony <- lmer(geometric.mean ~ timepoint + (1 | colony_id), data = oogen_size)

# Model without colony as a random effect
model_without_colony <- lm(geometric.mean ~ timepoint, data = oogen_size)

# run commparisons for AIC and BIC values
AIC(model_with_colony, model_without_colony)
BIC(model_with_colony, model_without_colony)
#model with colony as a random effect has both a lower AIC and BIC compared to the model without colony as a random effect, include colony_id

# Fit a mixed-effects model with 'patch' as a random effect
model.oogen <- lmer(geometric.mean ~ timepoint + (1 | colony_id), data = oogen_size)
qqPlot(residuals(model.oogen))

# Residuals are not normally distributed. Attempt with log transformation.  
model.oogen <-lmer(log(geometric.mean) ~ timepoint + (1 | colony_id), data = oogen_size)
qqPlot(residuals(model.oogen))

# Identify the outliers
outliers <- c(52,53)

# remove outliers
data_clean_oogen <- oogen_size[-outliers, ]

# Residuals are not normally distributed. Attempt with log transformation.  
model.oogen <-lmer(log(geometric.mean) ~ timepoint + (1 | colony_id), data = data_clean_oogen)
qqPlot(residuals(model.oogen))

# Get the summary of the model
summary(model.oogen)

# Perform Type III ANOVA
anova_result <- Anova(model.oogen, type = 3)
anova_result

# Post-hoc pairwise comparisons
pairwise_results.oogen <- emmeans(model.oogen, pairwise ~ timepoint, adjust = "tukey")

# Generate letters for significant differences
timepoint_cld_oogen <- cld(pairwise_results.oogen, Letters = letters)

# Get the maximum value of cells.cm2
max_oogen <- max(oogen_size$geometric.mean, na.rm = TRUE)  # Make sure to handle any missing values

# Make oogenesis plot
oogen <- oogen_size %>%
  ggplot(aes(x = timepoint, y = geometric.mean, fill = timepoint)) +
  geom_boxplot(color = "black", fill = "white", alpha = 0.2) +
  geom_point(pch = 21, size = 2, position = position_jitterdodge(0.2)) +
  ylim(0, 400) +
  geom_text(data = timepoint_cld_oogen, aes(label = .group, x = timepoint, y = 1.0 * max_oogen), vjust = -0.5, size = 14, color = "black") +
  ylab(expression(bold(paste("Oocyte Diameter (", mu, "m)")))) +  # Updated y-axis label
  theme_classic() +
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold", size = 30),
    axis.text = element_text(size = 30, color = "black"),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),  # Rotate x-axis labels
    strip.text.x = element_text(face = "italic", size = 30)); oogen

# Save the plot as an image
ggsave("histology/output/oogen.size.month.boxplot.png", oogen, width = 13, height = 8)


# save dataframe for downstream figure
write.csv(oogen_size, "gametogenesis_timeseries/gametogenesis_timeseries_analysis/Output/oogen_size_data.csv")
```


# Conduct univariate analyses and plot spermary size per month
```{r}
# Select for grand mean size dataframe
sperm_size <- abundance %>%
  dplyr::select(replicate_id, colony_id, timepoint, gonad.type, geometric.mean, Stage)

# Select just spermatocytes, remove any oocyte data, and filter out zero values
sperm_size <- sperm_size %>%
  filter(gonad.type == "spermatocyte") %>%
  filter(geometric.mean != 0)

# Count number of spermatocytes per timepoint for count on x-axis
sperm_size <- sperm_size %>%
  group_by(colony_id, timepoint) %>%
  mutate(count = if (all(is.na(geometric.mean) | geometric.mean == 0)) 0 else n()) %>%
  ungroup()

# Reorder the levels of the 'timepoint' variable from December to October
sperm_size$timepoint <- factor(sperm_size$timepoint, 
                               levels = c("December", "January", "February", "March", "April", "May", "June", "July", "August", "September", "October"),
                               ordered = TRUE)

# Calculate mean spermatocyte size per month per replicate (keep this step as needed for other purposes)
mean_size_month_sperm <- sperm_size %>%
  group_by(timepoint) %>%
  summarise(mean_size = mean(geometric.mean, na.rm = TRUE),
            std_error = sd(geometric.mean, na.rm = TRUE) / sqrt(n())) 

# Compare if colony_id as random effect improves model
# Model with colony as a random effect
model_with_colony <- lmer(geometric.mean ~ timepoint + (1 | colony_id), data = sperm_size)

# Model without colony as a random effect
model_without_colony <- lm(geometric.mean ~ timepoint, data = sperm_size)

# Run comparisons for AIC and BIC values
AIC(model_with_colony, model_without_colony)
BIC(model_with_colony, model_without_colony)
# Model with colony as a random effect has both a lower AIC and BIC compared to the model without colony as a random effect, include colony_id

# Fit a mixed-effects model with 'colony_id' as a random effect
model.sperm <- lmer(geometric.mean ~ timepoint + (1 | colony_id), data = sperm_size)
qqPlot(residuals(model.sperm))

# Residuals are not normally distributed. Attempt with log transformation.
model.sperm <- lmer(log(geometric.mean) ~ timepoint + (1 | colony_id), data = sperm_size)
qqPlot(residuals(model.sperm))

# Identify the outliers
outliers <- c(34,35)  # Adjust as necessary based on actual data

# Remove outliers
data_clean_sperm <- sperm_size[-outliers, ]

# Residuals are not normally distributed. Attempt with log transformation.
model.sperm <- lmer(log(geometric.mean) ~ timepoint + (1 | colony_id), data = data_clean_sperm)
qqPlot(residuals(model.sperm))

# Get the summary of the model
summary(model.sperm)

# Perform Type III ANOVA
anova_result_sperm <- Anova(model.sperm, type = 3)
anova_result_sperm

# Post-hoc pairwise comparisons
pairwise_results_sperm <- emmeans(model.sperm, pairwise ~ timepoint, adjust = "tukey")

# Generate letters for significant differences
timepoint_cld_sperm <- cld(pairwise_results_sperm, Letters = letters)

# Get the maximum value of geometric.mean
max_sperm <- max(sperm_size$geometric.mean, na.rm = TRUE)  # Make sure to handle any missing values

# Make spermatocyte plot
sperm <- sperm_size %>%
  ggplot(aes(x = timepoint, y = geometric.mean, fill = timepoint)) +
  geom_boxplot(color = "black", fill = "white", alpha = 0.2) +
  geom_point(pch = 21, size = 2, position = position_jitterdodge(0.2)) +
  ylim(0, 120) +
  geom_text(data = timepoint_cld_sperm, aes(label = .group, x = timepoint, y = 1.0 * max_sperm), vjust = -0.5, size = 14, color = "black") +
  ylab(expression(bold(paste("Spermary Diameter (", mu, "m)")))) +  # Updated label
  xlab("Timepoint") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.title = element_text(face = "bold", size = 30),
    axis.text = element_text(size = 30, color = "black"),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),  # Rotate x-axis labels
    strip.text.x = element_text(face = "italic", size = 30)); sperm

# Save the plot as an image
ggsave("histology/output/sperm.size.month.boxplot.png", sperm, width = 13, height = 8)

# save dataframe for downstream figure
write.csv(sperm_size, "gametogenesis_timeseries/gametogenesis_timeseries_analysis/Output/sperm_size_data.csv")

```

# Combine spermato and oogen size plots with patchwork
```{r}
# Combine plots using patchwork with a shared x-axis, tags, and centered axis labels
combined_plot <- (oogen / sperm) +
  plot_annotation(tag_levels = 'A') &
  theme(
    plot.tag = element_text(size = 30, face = "bold")  # Adjust tag text style here
  ) +
  plot_layout(guides = "collect")  # Collect the legends at the combined level

# Save the combined plot
ggsave("histology/output/combined_oocyte_sperm_size_plots.png", combined_plot, width = 20, height = 20)
```


# Calculate relative abundance change overtime with multinominal logistic regression
```{r}
# Use abundance clean dataframe for oocytes and spermaries
oocyte_data <- oocyte.abund.all
spermatocyte_data <- sperm_abun

# Use multinominal logistic regresssion to compare chnage in relative abundance overtime due to time as a predictor, categorical dependent variable and ordinal nature of stages
multinom_model <- multinom(Stage ~ timepoint, data = oogen_df_relative_abundance, weights = freq)
model_summary <- summary(multinom_model_with_colony)

#The multinom package does not include p-value calculation for the regression coefficients, so we calculate p-values using Wald tests (here z-tests).
# Extract coefficients and standard errors
coefficients <- model_summary$coefficients
std_errors <- model_summary$standard.errors

# Calculate z-scores
z_scores <- coefficients / std_errors

# Calculate two-tailed p-values
p_values <- 2 * (1 - pnorm(abs(z_scores)))

# Create a data frame with coefficients, standard errors, z-scores, and p-values
results <- data.frame(
  Coefficient = as.vector(coefficients),
  Std_Error = as.vector(std_errors),
  z_value = as.vector(z_scores),
  p_value = as.vector(p_values)
)

# Add row and column names
row.names(results) <- paste(rep(row.names(coefficients), each = ncol(coefficients)),
                            rep(colnames(coefficients), nrow(coefficients)),
                            sep = ":")

print(results)


```







